---
title: Model training
author: Marcelo Araya-Salas, PhD & Fabiola Chirino
date: '`r Sys.Date()`'
output:
  html_document:
    toc: yes
    toc_depth: 3
    code_folding: show
    df_print: kable
    toc_float:
      collapsed: no
      smooth_scroll: no
  pdf_document:
    toc: yes
    toc_depth: '3'
always_allow_html: true
subtitle: Vocal activity in A. lemur
fontsize: 12pt
editor_options:
  chunk_output_type: console
---

## Load packages
```{r packages, message = FALSE, warning = FALSE, eval = TRUE}

## add 'developer/' to packages to be installed from github
x <- c("devtools", "maRce10/warbleR", "bioacoustics", "pbapply", "Rraven",
       "parallel", "viridis", "ggplot2", "knitr", "kableExtra", 
       "maRce10/ohun", "DT", "formatR", "ranger")

aa <- lapply(x, function(y) {
  
  # get pakage name
  pkg <- strsplit(y, "/")[[1]]
  pkg <- pkg[length(pkg)]
  
  # check if installed, if not then install 
  if (!pkg %in% installed.packages()[,"Package"])  {

      if (grepl("/", y))  devtools::install_github(y, force = TRUE) else
    install.packages(y) 
    }

  # load package
  try(require(pkg, character.only = T), silent = T)
})

opts_knit$set(root.dir = "..")

```

```{r functions, eval = TRUE, echo = FALSE}

#warbleR_options(wav.path = "./data/raw/sound_files/10_kHz_5_min_cuts", wl = 300, parallel = parallel::detectCores() - 3, bp = "frange", fast = F, threshold = 15, ovlp = 20)

# warbleR_options(wav.path = "/media/m/Expansion/Projects/Ongoing/Agalychnis_lemur/sound_files/10_kHz_5_min_cuts", wl = 300, parallel = parallel::detectCores() - 3, bp = "frange", fast = F, threshold = 15, ovlp = 20)

# set evaluation false
opts_chunk$set(fig.width = 10, fig.height = 6, warning = FALSE, message = FALSE, tidy = TRUE)

# ggplot global setting
theme_set(theme_classic(base_size = 20))

```

# Downsample
```{r, eval = FALSE}

iw <- info_wavs()

table(iw$sample.rate)

# fix_wavs(samp.rate = 10)

```

# Create stratified 5 minute cuts
```{r create stratified 5 minute cuts, eval = FALSE}

# path with waves
#wav_path <- "/Volumes/CHIRINO/Grabaciones SM4 2019/"

wav_path <- paste0("/media/m/Expansion/Projects/Ongoing/Agalychnis_lemur",
                  "/sound_files/10_kHz_5_min_cuts/converted_sound_files/")

# list of wave files
wavs <- list.files(path = wav_path, pattern = "\\.wav$", ignore.case = TRUE,
                   recursive = TRUE, full.names = TRUE)

# data frame with waves and metadata
wav_df <- data.frame(path = dirname(wavs), file =  basename(wavs))
wav_df$site <- ifelse(grepl("sukia", wav_df$file, ignore.case = TRUE), 
                      "Sukia", "Chimu") 
sm_metadata <- seewave::songmeter(wav_df$file)

wav_df <- cbind(wav_df, sm_metadata)

head(wav_df)

str(wav_df)

par(mfrow = c(2, 1))
hist(wav_df$time[wav_df$site == "Sukia"], breaks = 100, 
     col = adjustcolor("green", alpha.f = 0.4), main = "Sukia")
hist(wav_df$time[wav_df$site == "Chimu"], breaks = 100, 
     col = adjustcolor("blue", alpha.f = 0.4), main = "Chimu")

wav_df <- wav_df[order(wav_df$time), ]

wav_df$day_count <- as.vector(round((wav_df$time - min(wav_df$time)) /
                                      (3600 * 24), 0) + 1)

wav_df$week <- cut(x = wav_df$day_count, breaks = seq(1, max(wav_df$day_count) +
                                                        7, by = 7), include.lowest = TRUE)

wav_df <- droplevels(wav_df)

wav_df$period <- 4
wav_df$period[wav_df$hour %in% 17:19] <- 1
wav_df$period[wav_df$hour %in% 20:22] <- 2
wav_df$period[wav_df$hour %in% c(23, 0, 1)] <- 3

# create name combination
wav_df$site.week.period <- paste(wav_df$site, wav_df$week, wav_df$period, sep = "-")

# remove LAguna chimu
wav_df <- wav_df[grep("LAguna Chimu", wav_df$path, invert = TRUE), ]

set.seed(1)
sample_wav_l <- lapply(unique(wav_df$site.week.period), function(x){
  
  X <- wav_df[wav_df$site.week.period == x, ]
  
  if(nrow(X) > 2)
    X <- X[sample(1:nrow(X), size = 2), ]
  
  return(X)
    
}) 

sample_wav_df <- do.call(rbind, sample_wav_l)

write.csv(sample_wav_df, 
  file.path("/Volumes/CHIRINO/Proyecto Lemur/INV_2021/5_min_cut_lemur/",
                         "5_min_cut_metadata.csv"), row.names = FALSE)

sample_wav_df$cut.name <- gsub(".wav",  "_5min_cut.wav", sample_wav_df$file)

library(pbapply)

out <- pblapply(1:nrow(sample_wav_df), function(x) {

  # read first 5 mins
  wv <- try(readWave(file.path(sample_wav_df$path[x], sample_wav_df$file[x]), 
                     from = 0, to = 300, units = "seconds"), silent = TRUE)
  
  if (class(wv) == "Wave")
  # save cut
  writeWave(wv, filename = 
              file.path("/Volumes/CHIRINO/Proyecto Lemur/INV_2021/5_min_cut_lemur/", 
                        sample_wav_df$cut.name[x])) 

  if (class(wv) == "Wave")
  return(NULL) else return(x)
})
```

## Cheking manual selections
```{r Cheking manual selections, eval = FALSE}

# list all selection tables
sl_tbs <- list.files(paste0("~/Dropbox/estudiantes/Fabiola/proyecto_lemur/data", 
              "/processed/selection_tables/all_selec_tables_lemur/"))

# remove empty selection table
sl_tbs <- sl_tbs[sl_tbs != "LAGCHIMU_20200919_201500.Table.2.selections.txt"]

# import all selections
sls <- imp_raven(path = paste0("~/Dropbox/estudiantes/Fabiola/proyecto_lemur/data", 
              "/processed/selection_tables/all_selec_tables_lemur/"), warbler.format = TRUE, all.data = TRUE, files = sl_tbs)

# relabel selec column
sls$selec <- 1:nrow(sls)

# check duration
sls$duration  <- sls$end - sls$start
unique(sls$sound.files[sls$duration > 1])

# remove temporarily but must be fixed on the original individual selection tables
sls <- sls[sls$duration < 1, ]

# number of sound files and selection files should be the same
length(unique(sls$sound.files)) == length(unique(sls$selec.file))

# find files with more than 1 selection
# tab <- table(sls$sound.files[!duplicated(sls$selec.file)]) 
# tab[tab > 1]
# unique(sls$selec.file[sls$sound.files == "LAGCHIMU_20190709_200000_5min_cut.wav"])

# check selections
cs <- check_sels(X = sls, pb = FALSE)
table(cs$check.res[cs$check.res != "OK"])

# those with errors were empty selections (single points)
sls <- sls[cs$check.res == "OK", ]

# check for duplicates (if any sound files was anottated more than once)
sls <- overlapping_sels(sls, indx.row = TRUE, pb = FALSE)
sls <- sls[!duplicated(sls$indx.row, incomparables = NA), ]

# should be 4 in "SUKIA_20200207_220000_5min_cut.wav"
table(sls$sound.files[!is.na(sls$indx.row)])

# export raven multiple sound file selection table
exp_raven(sls, path = "./data/processed/selection_tables/", sound.file.path = .Options$warbleR$path, file.name = "temporary_doublechecking_all_selections")


# fix path for Fabis laptop
fix_path(path =  "./data/processed/selection_tables/",
         new.begin.path = paste0("/Users/fabiolachirino/Dropbox/Fabiola/proyecto_lemur",
                  "/data/raw/sound_files/10_kHz_5_min_cuts/"), 
         sound.file.col = "Begin File")


# create sound file labels
sf.abbr <- gsub("0000_5min_cut.wav", "", sls$sound.files)
sf.abbr <- gsub("SUKIA_20", "S", sf.abbr)
sls$sf.abbr <- gsub("LAGCHIMU_20", "L", sf.abbr)


sls_est <- selection_table(sls, path = "/media/m/CHIRINO/Proyecto Lemur/INV_2021/5_min_cut_data/converted_sound_files", extended = TRUE)

saveRDS(sls_est, "./data/processed/extended_selection_table_manual_annotations_lemur.RDS")

```

```{r copy files for testing detection methods python, eval = FALSE}

traing_dat <- imp_raven(path = "./data/processed/selection_tables/", files = "temporary_doublechecking_all_selections.txt")

train_files <- unique(traing_dat$`Begin File`)

fc <- file.copy(from = file.path("/media/m/Expansion/Projects/Ongoing/Agalychnis_lemur",
          "sound_files/10_kHz_5_min_cuts/", train_files), to = file.path("/media/m/Expansion/Projects/Ongoing/Agalychnis_lemur",
                                                                         "/sound_files/training_sound_files/", train_files))

sls$tags <- "lemur"
exp_raven(sls[, c("sound.files", "selec", "start", "end", "bottom.freq", "top.freq", "tags")], path = paste0("/media/m/Expansion/Projects/Ongoing/Agalychnis_lemur",
                                                  "/sound_files/training_sound_files/"), 
          sound.file.path = .Options$warbleR$path, 
          file.name = "annotations_raven_format")

fix_path(path = paste0("/media/m/Expansion/Projects/Ongoing/Agalychnis_lemur",
                                                  "/sound_files/training_sound_files/"), 
         new.begin.path = "", sound.file.col = "Begin File")
fix_path(path = paste0("/media/m/Expansion/Projects/Ongoing/Agalychnis_lemur",
                                                  "/sound_files/training_sound_files/"),
         new.begin.path = "", sound.file.col = "Begin Path")
```


## Create catalogs
```{r Create catalogs, eval = FALSE}

sls_est <- readRDS("./data/processed/extended_selection_table_manual_annotations_lemur.RDS")

# create catalogs
catalog(X = sls_est, flim = c(1, 3.5), nrow = 12, ncol = 15, ovlp = 70, height = 15, width = 23, same.time.scale = TRUE, mar = 0.005, wl = 512, gr = FALSE,
        spec.mar = 0.4, lab.mar = 0.8, rm.axes = TRUE, by.row = TRUE, 
        box = TRUE, labels = c("sf.abbr", "selec"), fast.spec = TRUE, 
        pal = viridis, parallel = 10)

# move images to figures folder
move_imgs(from = .Options$warbleR$path, to = "./figures", overwrite = TRUE)


# catalog for paper
catalog(X = sls_est[1:180,], flim = c(1, 3.5), nrow = 12, ncol = 15, ovlp = 90, 
        height = 15, width = 23, same.time.scale = TRUE, mar = 0.1, wl = 512,
        gr = FALSE, spec.mar = 0, lab.mar = 0.0001, rm.axes = TRUE, 
        by.row = TRUE, box = FALSE, fast.spec = FALSE, pal = viridis, 
        parallel = 1, img.suffix = "no_margins", collevels = seq(-115, 0, 5))

# move images to figures folder
move_imgs(from = .Options$warbleR$path, to = "./figures", overwrite = TRUE)


```

## Measure frequency range parameters

To figure out variation in signal structure aiming to define templates
```{r Measure frequency range parameters, eval = FALSE, results='asis'}
# measure spectrographic parameters
spectral_parameters <- spectro_analysis(sls, bp = "frange", fast = TRUE, 
                                        ovlp = 70, pb = FALSE)

saveRDS(spectral_parameters, "./data/processed/acoustic_parameters_and_selections.RDS")

```


```{r plot frequency range parameters, eval = TRUE, results='asis'}

spectral_parameters <- readRDS("./data/processed/acoustic_parameters_and_selections.RDS")

sls <- imp_raven(path = "./data/processed/selection_tables/", files = "temporary_doublechecking_all_selections.txt", pb = FALSE, warbler.format = TRUE)

# measure signal_2_noise ratio
# sls <- signal_2_noise(sls, mar = 0.1, before = TRUE, pb = FALSE, bp = "frange")
pca <- prcomp(spectral_parameters[, 2:27], scale. = TRUE)


# extrac SNR and peak freq
sls$meanpeakf <- spectral_parameters$meanpeakf
sls$duration <- spectral_parameters$duration
sls$sp.ent <- spectral_parameters$sp.ent
sls$PC1 <- pca$x[, 1]


# stck_parameters$PC1
stck_parameters <- stack(sls[, c("duration", "meanpeakf", "sp.ent", "PC1")])

ggplot(stck_parameters, aes(x = values)) +
  geom_histogram(fill = viridis(10, alpha = 0.9)[2]) +
  facet_wrap(~ ind, scales = "free")

spectral_parameters <- spectral_parameters[!is.infinite(spectral_parameters$SNR), ]

freq_variation <- sapply(c("duration", "meanpeakf", "sp.ent", "PC1"), function(x) 
  data.frame(mean = mean(sls[, x]), min = min(sls[, x]), max = max(sls[, x]), 
             sd = sd(sls[, x]), CI.2.5 = quantile(sls[, x], 0.025), 
             CI.97.5 = quantile(sls[, x], 0.975)))

freq_variation <- do.call(rbind, lapply(data.frame(freq_variation), function(x) round(unlist(x), 3)))

# print table as kable  
kb <-kable(freq_variation, row.names = TRUE, digits = 3) 
  
kb <- kable_styling(kb, bootstrap_options = c("striped", "hover", "condensed", "responsive"))
  
print(kb)

```

## Template-based detection
```{r  template creation, eval = FALSE}


# find template close to mean peak freq
mean_peak_indx <- which.min(abs(sls$meanpeakf - freq_variation[rownames(freq_variation) == "meanpeakf", 1]))

# find template close to mean duration
mean_duration_indx <- which.min(abs(sls$duration - freq_variation[rownames(freq_variation) == "duration", 1]))

# find template close to mean PC1
mean_pc1_indx <- which.min(abs(sls$PC1 - freq_variation[rownames(freq_variation) == "PC1", 1]))

# put both templates together
templates <- sls[c(mean_peak_indx, mean_duration_indx, mean_pc1_indx), ]

# label templates
templates$template.type <- c("mean.peak.freq", "mean.duration", "mean.PC1")

#  create ext. selection table
templates_est <- selection_table(templates, extended = TRUE, 
                                 confirm.extended = FALSE, fix.selec = TRUE)

templates_est <- rename_est_waves(templates_est, new.sound.files = templates$template.type)

catalog(templates_est, nrow = 2, ncol = 2, ovlp = 70, height = 10, width = 14,
        same.time.scale = TRUE, mar = 1, wl = 512, gr = FALSE, spec.mar = 0.6,
        lab.mar = 1, rm.axes = FALSE, by.row = TRUE, box = TRUE, 
        labels = "sound.files", fast.spec = FALSE, pal = viridis, 
        flim = c(1, 3.5), img.suffix = "templates", collevels = seq(-100, 0, 5))

te2 <- te <- templates_est[rep(2:4, 2), ]
# te2 <- rename_est_waves(te2, new.sound.files = letters[1:3])
te2$selec <- 2
te3 <- rbind(te, te2)

catalog(templates_est, nrow = 2, ncol = 3, ovlp = 90, height = 10, width = 14, same.time.scale = TRUE, mar = 1, wl = 512, gr = FALSE, spec.mar = 0.6, 
        lab.mar = 2, rm.axes = FALSE, by.row = TRUE, box = FALSE, 
        labels = "sound.files", fast.spec = FALSE, pal = viridis, 
        flim = c(1, 3.5), img.suffix = "templates", collevels = seq(-100, 0, 5))

# # move images to figures folder
move_imgs(from = .Options$warbleR$path, to = "./figures", overwrite = TRUE)

# save est
saveRDS(templates_est, "./data/processed/mean_acoustic_parameter_templates_est.RDS")

```

## Detection perfomance

Using 4 templates:

  - mean_peakf: the call with the peak frequency closest to the mean of the population 

```{r optimization results get correlations, eval = FALSE}
templates_est <- 
  readRDS("./data/processed/mean_acoustic_parameter_templates_est.RDS")

paral <- 15

# use mean peak freq sels as template, fourier spectrograms
corr_templ <- template_correlator(templates = templates_est, 
                                  files = unique(sls$sound.files), 
                                  path = .Options$warbleR$path, 
                                  parallel = paral, hop.size = 11.6, ovlp = 70)
  
saveRDS(corr_templ, "./data/processed/template_correlations_4_templates.RDS")
```


```{r optimization run, eval = FALSE}

corr_templ <- readRDS("./data/processed/template_correlations_4_templates.RDS")

optimize_fourier_detec <- optimize_template_detector(reference = sls, template.correlations = corr_templ, threshold = seq(0.05, 0.5, 0.01), 
  parallel = 1, by.sound.file = TRUE, pb =  TRUE)

optimize_fourier_detec$recall <- optimize_fourier_detec$sensitivity
optimize_fourier_detec$precision <- optimize_fourier_detec$specificity
optimize_fourier_detec$sensitivity <- optimize_fourier_detec$specificity <- NULL
optimize_fourier_detec$f1.score <- optimize_fourier_detec$recall * optimize_fourier_detec$precision

saveRDS(optimize_fourier_detec, "./data/processed/optimization_results_4_templates.RDS")

```

```{r read detection optimization results, eval = FALSE, results='asis'}
optimize_fourier_detec <- readRDS("./data/processed/optimization_results_4_templates.RDS")

optimize_fourier_detec$templates <- gsub("-1", "",optimize_fourier_detec$templates)

optimize_fourier_detec$`overlap.to.true.positives` <- optimize_fourier_detec$proportional.overlap.true.positives

sd_tab <- summarize_diagnostic(optimize_fourier_detec)

saveRDS(sd_tab, "./data/processed/summary_optimization_results_4_templates.RDS")
```

```{r read detection optimization results 2, eval = TRUE, results='asis'}

sd_tab <- readRDS("./data/processed/optimization_results_4_templates.RDS")

sd_tab <- sd_tab[grep("SNR", sd_tab$templates, invert = TRUE),]

# sd_tab$f1.score <- sd_tab$rec * sd_tab$specificity

agg_sd <- aggregate(cbind(recall, precision, f1.score) ~ threshold + 
                      templates, data = sd_tab, mean)

ggplot(agg_sd, aes(x = threshold, y = recall, group = templates, 
                   color = templates)) +
  geom_line() +
  geom_point() +
  scale_color_viridis_d(end = 1) +
  theme_classic() 

ggplot(agg_sd, aes(x = threshold, y = precision, group = templates, 
                   color = templates)) +
  geom_line() +
  geom_point() +
  scale_color_viridis_d(end = 0.8) +
  theme_classic() 

ggplot(agg_sd, aes(y = precision, x = recall, group = templates, 
                   color = templates)) +
  geom_line() +
  geom_point() +
  scale_color_viridis_d(end = 1) +
  theme_classic() 

ggplot(agg_sd, aes(x = threshold, y = f1.score, group = templates, 
                   color = templates)) +
  geom_line() +
  geom_point() +
  scale_color_viridis_d(end = 0.8) +
  theme_classic() 


stck_agg_sd <- cbind(agg_sd[, 1:2], stack(agg_sd[, 3:5]))

stck_agg_sd$ind <- factor(stck_agg_sd$ind, labels = c("Recall", "Precision", 
                                                      "F1 score"))

# ggf1 <-
  ggplot(stck_agg_sd, aes(x = threshold, y = values, group = templates, 
                          color = templates)) +
  geom_line() +
  geom_point() +
  labs(x = "Correlation threshold", color = "Template", y = "") +
  facet_wrap(~ind, nrow = 1, scales = "free_y") + 
  scale_color_viridis_d(end = 0.8, 
                        labels = c("Mean duration", "Mean PC1", 
                                   "Mean peak frequency"), alpha = 0.5) +
  theme_classic() 

# ggsave("./data/processed/multipanel_perfomance_indices.jpeg", width = 9, height = 4, dpi = 300)
  
# print dynamic table
oa_DT <- datatable(sd_tab, editable = list(
  target = 'row'
), rownames = FALSE, style = "bootstrap",  filter = 'top', 
options = list(
  pageLength = 100, autoWidth = TRUE, dom = 'ft'
), autoHideNavigation = TRUE, escape = FALSE)

formatRound(table = oa_DT, columns = sapply(sd_tab, is.numeric), 3)

  
```


# Template fourier mean.PC1 threshold 0.43

Choose 'fourier mean.PC1' template with a threshold of 0.43 due to a good sensitivity and relatively low number of false positives

```{r, eval = FALSE}

templates_est <- readRDS("./data/processed/mean_acoustic_parameter_templates_est.RDS")

corr_templ <- template_correlator(templates = templates_est[templates_est$sound.files == "mean.PC1", , drop = FALSE], 
                                  files = unique(sls$sound.files), 
                                  path = .Options$warbleR$path, parallel = 10,
                                  hop.size = 11.6, ovlp = 70)

detec <- template_detector(template.correlations = corr_templ,
                           threshold = 0.43, parallel = 10, pb =  TRUE)

diagnostic <- diagnose_detection(reference = sls, detection = detec, parallel = 10,  
                                 by.sound.file = TRUE)

summarize_diagnostic(diagnostic)

saveRDS(diagnostic,
        "./data/processed/diagnostic_mean_pc1_0.43_threshold_diagnostic.RDS")

lab_detec <- label_detection(reference = sls, detection = detec, parallel = 10)

saveRDS(lab_detec, 
        "./data/processed/label_detection_mean_pc1_0.43_threshold.RDS")

```

```{r filter detections, eval = FALSE}

lab_detec <- 
  readRDS("./data/processed/label_detection_mean_pc1_0.43_threshold.RDS")

filter_lab_detec <- filter_detection(detection = lab_detec, by = "scores")

saveRDS(filter_lab_detec,
        "./data/processed/filtered_detection_mean_pc1_0.43_threshold.RDS")

diagnose_model <- diagnose_detection(reference = sls, filter_lab_detec, 
                                     parallel = 10)


diagnose_model$f1.score <- 
  diagnose_model$specificity * diagnose_model$sensitivity

saveRDS(diagnose_model, "./data/processed/diagnostics_detection_mean_pc1_0.43_threshold.RDS")

```

```{r diagnostic, eval = TRUE}

diag <- 
  readRDS("./data/processed/diagnostics_detection_mean_pc1_0.43_threshold.RDS")

# print dynamic table
oa_DT <- datatable(diag, editable = list(
  target = 'row'
), rownames = FALSE, style = "bootstrap",  filter = 'top', 
options = list(
  pageLength = 100, autoWidth = TRUE, dom = 'ft'
), autoHideNavigation = TRUE, escape = FALSE)

formatRound(table = oa_DT, columns = sapply(diag, is.numeric), 3)

```

## Random forest results
```{r measure acoustic parameters and run random forest, eval = FALSE}

filter_lab_detec <- readRDS("./data/processed/filtered_detection_mean_pc1_0.43_threshold.RDS")

# measure spectrographic parameters
spectral_parameters <- spectro_analysis(filter_lab_detec, bp = c(1, 3.5), 
                                        fast = TRUE, ovlp = 70, parallel = 10)

# mfccs <- mfcc_stats(X = lab_detec, bp = c(1, 3.5), ovlp = 70, parallel = 10)
# 
# na_rows <- unique(unlist(sapply(mfccs, function(x) which(is.na(x)))))
# 
# lab_detec <- lab_detec[-na_rows, ]
# spectral_parameters <- spectral_parameters[-na_rows, ]
# mfccs <- mfccs[-na_rows, ]

spectral_parameters$class <- filter_lab_detec$detection.class

# spectral_parameters <- data.frame(spectral_parameters, mfccs[, !names(spectral_parameters) %in% c("sound.files", "selec")])

spectral_parameters$class[spectral_parameters$class != "false.positive"] <-
  "true.positive"

# make it a factor for ranger to work 
spectral_parameters$class <- as.factor(spectral_parameters$class)
  
# run RF model spectral and cepstral parameters
  rfm <-
    ranger(
      class ~ .,
      data = spectral_parameters[, !names(spectral_parameters) %in% c("sound.files", "selec")],
      num.trees = 10000,
      importance = "impurity",
      seed = 10
    )
  
saveRDS(list(rfm = rfm, spectral_parameters = spectral_parameters,
             filter_lab_detec = filter_lab_detec),
paste0("./data/processed/data_and_model_random_forest_0.43_",
       "threshold_only_spectro_parameters.RDS"))  

```

```{r random forest results, eval = TRUE}

# attach(readRDS("./data/processed/data_and_model_random_forest_0.43_threshold.RDS"))


attach(readRDS(paste0("./data/processed/data_and_model_random_forest_0.43_",
                      
       "threshold_only_spectro_parameters.RDS")))

rfm

```

Diagnostic after random forest classification:
```{r random forest diagnosticsi, eval = TRUE}

filter_lab_detec$pred.class <- rfm$predictions

positive_detec <- filter_lab_detec[filter_lab_detec$pred.class == 
                                     "true.positive", ]



#save for trying stats
# write.csv(positive_detec, "./data/processed/true_positive_detections_on_training_recordings.csv", row.names = FALSE)

# exp_raven(positive_detec, file.name = "./data/processed/true_positive_detections_on_training_recordings", sound.file.path = .Options$warbleR$path)

# fix path for Fabis laptop
#fix_path(path =  "./data/processed/",new.begin.path = "/Users/fabiolachirino/Dropbox/Fabiola/proyecto_lemur/data/raw/sound_files/10_kHz_5_min_cuts/", sound.file.col = "Begin File")

temp_detec <- positive_detec
temp_detec$detection.class <- "true.positive"

diag <- diagnose_detection(reference = sls, detection = temp_detec, pb = FALSE)

# print dynamic table
oa_DT <- datatable(diag, editable = list(
  target = 'row'
), rownames = FALSE, style = "bootstrap",  filter = 'top', 
options = list(
  pageLength = 100, autoWidth = TRUE, dom = 'ft'
), autoHideNavigation = TRUE, escape = FALSE)

formatRound(table = oa_DT, columns = sapply(diag, is.numeric), 3)

```

 Black line = 1:1
 gray line = model slope

```{r random forest results plots, eval = TRUE}

obs_count <- tapply(sls$sound.files, sls$sound.files, length)
pred_count <- tapply(positive_detec$sound.files, positive_detec$sound.files, length)

int_columns <- intersect(names(obs_count), names(pred_count))
obs_count <- obs_count[names(obs_count) %in% int_columns]
pred_count <- pred_count[names(pred_count) %in% int_columns]
pred_count <- pred_count[order(names(pred_count))]
obs_count <- obs_count[order(names(obs_count))]

df <- data.frame(sound.files = names(obs_count), observed = obs_count, predicted = pred_count)

ggplot(df, aes(x = observed, y = predicted)) +
  geom_point(color = viridis(10, alpha = 0.4)[2], size = 3) +
  geom_abline(slope = 1, intercept = 0) +
  annotate("text", x = 50, y = 150, 
           label = paste("r =", round(cor(obs_count, pred_count), 3)), size = 8) +
  geom_smooth(method = "lm", se = FALSE, col = "gray") +
  theme_classic(base_size = 20)

(lm(pred_count ~ obs_count))

```

Removing outlier
```{r random forest results plots 2, eval = TRUE}
ggplot(df[df$observed < 100, ], aes(x = observed, y = predicted)) +
  geom_point(color = viridis(10, alpha = 0.4)[2], size = 3) +
  geom_abline(slope = 1, intercept = 0) +
  # annotate("text", x = 50, y = 70, 
  #          label = paste("r =", round(cor(obs_count[obs_count < 100],
  #                       pred_count[obs_count < 100]), 3)), size = 8) + 
  geom_smooth(method = "lm", se = FALSE, col = "gray") +
  theme_classic(base_size = 20)


(lm(pred_count[obs_count < 100] ~ obs_count[obs_count < 100]))

```

```{r random forest results plots 2 for presentation, eval = TRUE}
ggplot(df[df$observed < 100, ], aes(x = observed, y = predicted)) +
  geom_point(color = alpha("#F2622E", 0.6), size = 3) +
  geom_abline(slope = 1, intercept = 0) +
  annotate("text", x = 35, y = 80, 
           label = paste("r =", round(cor(obs_count[obs_count < 100],
                        pred_count[obs_count < 100]), 3)), size = 8) + 
  geom_smooth(method = "lm", se = FALSE, col = "gray") +
  theme_classic(base_size = 20) + labs(x = "Cantos observados", y = "Cantos detectados")


(lm(pred_count[obs_count < 100] ~ obs_count[obs_count < 100]))

```

# Template fourier mean.PC1 threshold 0.38

Choose 'fourier mean.PC1' template with a threshold of 0.43 due to a good sensitivity and relatively low number of false positives

```{r 0.38, eval = FALSE}
templates_est <- readRDS(paste0("./data/processed/mean_acoustic_", "parameter_templates_est.RDS"))

corr_templ <- template_correlator(templates =
                                    templates_est[templates_est$sound.files ==
                                                    "mean.PC1", , drop = FALSE],
                                  files = unique(sls$sound.files), 
                                  path = .Options$warbleR$path, 
                                  parallel = 10, hop.size = 11.6, ovlp = 70)

detec <- template_detector(template.correlations = corr_templ, threshold = 0.38,
                           parallel = 10, pb =  TRUE)

diagnostic <- diagnose_detection(reference = sls, detection = detec,
                                 parallel = 10,  by.sound.file = TRUE)

summarize_diagnostic(diagnostic)

saveRDS(diagnostic, 
        "./data/processed/diagnostic_mean_pc1_0.38_threshold_diagnostic.RDS")

lab_detec <- label_detection(reference = sls, detection = detec, parallel = 10)

saveRDS(lab_detec, "./data/processed/label_detection_mean_pc1_0.38_threshold.RDS")
```

```{r filter detections 0.38, eval = FALSE}

lab_detec <-
  readRDS("./data/processed/label_detection_mean_pc1_0.38_threshold.RDS")

filter_lab_detec <- filter_detection(lab_detec, parallel = 10)

saveRDS(filter_lab_detec, "./data/processed/filtered_detection_mean_pc1_0.38_threshold.RDS")

diagnose_model <- diagnose_detection(reference = sls, 
                                     filter_lab_detec, parallel = 10)

saveRDS(diagnose_model, "./data/processed/diagnostics_detection_mean_pc1_0.38_threshold.RDS")

```


```{r diagnostic 0.38, eval = TRUE}

diag <- 
  readRDS("./data/processed/diagnostics_detection_mean_pc1_0.38_threshold.RDS")

# print dynamic table
oa_DT <- datatable(diag, editable = list(
  target = 'row'
), rownames = FALSE, style = "bootstrap",  filter = 'top', 
options = list(
  pageLength = 100, autoWidth = TRUE, dom = 'ft'
), autoHideNavigation = TRUE, escape = FALSE)

formatRound(table = oa_DT, columns = sapply(diag, is.numeric), 3)

```

## Random forest results
```{r measure acoustic parameters and run random forest 0.38, eval = FALSE}
lab_detec <-
  readRDS("./data/processed/label_detection_mean_pc1_0.38_threshold.RDS")

# measure spectrographic parameters
spectral_parameters <- spectro_analysis(lab_detec, bp = c(1, 3.5), fast = TRUE,
                                        ovlp = 70, parallel = 10)

# mfccs <- mfcc_stats(X = lab_detec, bp = c(1, 3.5), ovlp = 70, parallel = 10)
# 
# na_rows <- unique(unlist(sapply(mfccs, function(x) which(is.na(x)))))
# 
# lab_detec <- lab_detec[-na_rows, ]
# spectral_parameters <- spectral_parameters[-na_rows, ]
# mfccs <- mfccs[-na_rows, ]

spectral_parameters$class <- lab_detec$detection.class

# spectral_parameters <- data.frame(spectral_parameters, mfccs[, !names(spectral_parameters) %in% c("sound.files", "selec")])

spectral_parameters$class[spectral_parameters$class != "false.positive"] <-
  "true.positive"

# make it a factor for ranger to work 
spectral_parameters$class <- as.factor(spectral_parameters$class)
  
  # run RF model spectral and cepstral parameters
  rfm <-
    ranger(
      class ~ .,
      data = spectral_parameters[, !names(spectral_parameters) %in% c("sound.files", "selec")],
      num.trees = 10000,
      importance = "impurity",
      seed = 10
    )
  
saveRDS(list(rfm = rfm, spectral_parameters = spectral_parameters, 
             lab_detec_0.38 = lab_detec),
"./data/processed/data_and_model_random_forest_0.38_threshold_only_spectro_parameters.RDS")  
```

```{r random forest results 0.38, eval = TRUE}
attach(readRDS("./data/processed/data_and_model_random_forest_0.38_threshold_only_spectro_parameters.RDS"))

rfm
```

Diagnostic after random forest classification:
```{r random forest diagnostic 0.38, eval = TRUE}
lab_detec <- lab_detec_0.38
# table(lab_detec$detection.class)
lab_detec$pred.class <- rfm$predictions

positive_detec <- lab_detec[lab_detec$pred.class == "true.positive", ]


temp_detec <- positive_detec
temp_detec$detection.class <- "true.positive"

diag <- diagnose_detection(reference = sls, detection = temp_detec, pb = FALSE)

# print dynamic table
oa_DT <- datatable(diag, editable = list(
  target = 'row'
), rownames = FALSE, style = "bootstrap",  filter = 'top',
options = list(
  pageLength = 100, autoWidth = TRUE, dom = 'ft'
), autoHideNavigation = TRUE, escape = FALSE)

formatRound(table = oa_DT, columns = sapply(diag, is.numeric), 3)
```

 Black line = 1:1
 gray line = model slope
```{r random forest results plots 0.38, eval = TRUE}

obs_count <- tapply(sls$sound.files, sls$sound.files, length)
pred_count <- tapply(positive_detec$sound.files, positive_detec$sound.files, length)

int_columns <- intersect(names(obs_count), names(pred_count))
obs_count <- obs_count[names(obs_count) %in% int_columns]
pred_count <- pred_count[names(pred_count) %in% int_columns]
pred_count <- pred_count[order(names(pred_count))]
obs_count <- obs_count[order(names(obs_count))]

df <- data.frame(sound.files = names(obs_count), observed = obs_count, predicted = pred_count)

ggplot(df, aes(x = observed, y = predicted)) +
  geom_point(color = viridis(10, alpha = 0.4)[2], size = 3) +
  geom_abline(slope = 1, intercept = 0) +
  annotate("text", x = 50, y = 150, label = 
             paste("r =", round(cor(obs_count, pred_count), 3)), size = 8) + 
  geom_smooth(method = "lm", se = FALSE, col = "gray") +
  theme_classic(base_size = 20)

(lm(pred_count ~ obs_count))

```

Removing outlier
```{r random forest results plots 2 0.38, eval = TRUE}
ggplot(df[df$observed < 100, ], aes(x = observed, y = predicted)) +
  geom_point(color = viridis(10, alpha = 0.4)[2], size = 3) +
  geom_abline(slope = 1, intercept = 0) +
  annotate("text", x = 50, y = 150, 
           label = paste("r =", round(cor(obs_count[obs_count < 100],
                                          pred_count[obs_count < 100]), 3)), 
           size = 8) + 
  geom_smooth(method = "lm", se = FALSE, col = "gray") +
  theme_classic(base_size = 20)


(lm(pred_count[obs_count < 100] ~ obs_count[obs_count < 100]))

```


# Graphs for presentations

```{r}
library(ggplot2)
library(cowplot)

n <- 90000
sd <- 0.3
set.seed(1)
x <- rnorm(n, mean = 0.1, sd = sd)
set.seed(2)
y <- rnorm(n, mean = 0.9, sd = sd)
umbral <- 0.5
df <- data.frame(tipo = rep(c("Otros\nsonidos", "Cantos\nde lemur"), each = n), vals = c(x, y))

df$tipo <- factor(df$tipo, levels = c("Otros\nsonidos", "Cantos\nde lemur"))

df$vals <- (df$vals + abs(min(df$vals)))/ max(df$vals + abs(min(df$vals)))

fill_values <- c("#AFBF35","#F2622E")
fill_values <- adjustcolor(fill_values, alpha.f = 0.6)

gg1 <-
  ggplot(df, aes(x = vals, fill = tipo)) +   
  geom_density() +
 scale_fill_manual(values = fill_values) +
 labs(x = "Umbral de correlación", y = "", fill = "") +
  geom_vline(xintercept = umbral, lty = 2) +
  theme_classic(base_size = 25) +
  theme(legend.position = c(.9, .9)) 

agg <- aggregate(vals ~ tipo, df[df$vals > umbral, ], length)
agg$tipo <-  factor(agg$tipo, levels = c("Otros\nsonidos", "Cantos\nde lemur"))

gg2 <- ggplot(agg, aes(x = tipo, y = vals, fill = tipo)) +
  geom_col(col = "black") + 
  scale_fill_manual(values = fill_values, guide = 'none') +
  theme_classic(base_size = 25) + 
  labs(x = "", y = "Proporción") + ylim(c(0, n)) +
  theme(axis.text.y=element_blank())

cowplot::plot_grid(gg1, gg2, nrow = 1, rel_widths = c(1.7, 1))

# cowplot::ggsave2(plot = ggcp, filename = "./output/threshold_0.3.tiff", width = 15, height = 5)
```


---

<font size="4">Session information</font>

```{r session info, echo=F, eval = TRUE}

sessionInfo()

```
