---
title: "Size_effect_lemur_analysis"
author: "Fabiola Chirino"
date: "2022-12-05"
output: html_document
---

Temperature model: 
```{r}
# Standard deviation transformation: 

exp(0.307) # effect 
exp(0.257) # CI_low 
exp(0.357) #CI_high

# sd & unit transformation: 

(exp(0.307) - 1) / 2.12 # sd without any transformation

(exp(0.307) - 1) / 2.12 + 1

```
Temperature model (24h) (Version 2): 
```{r}
# Standard deviation transformation: 

exp(0.312) # effect 
exp(0.261) # CI_low 
exp(0.361) #CI_high

# sd & unit transformation: 

(exp(0.312) - 1) / 2.12 # sd without any transformation

(exp(0.312) - 1) / 2.12 + 1
```
Temperature model (48h) (Version 2): 
```{r}
# Standard deviation transformation: 

exp(0.305) # effect 
exp(0.255) # CI_low 
exp(0.355) #CI_high

# sd & unit transformation: 

(exp(0.305) - 1) / 2.12 # sd without any transformation

(exp(0.305) - 1) / 2.12 + 1
```

Previous Temperature (Version 2): 
```{r}
# Standard deviation transformation: 

exp(-0.002) # effect 
exp(-0.048) # CI_low 
exp(0.044) #CI_high

# sd & unit transformation: 

(exp(-0.002) - 1) / 0.988 # sd without any transformation

(exp(-0.002) - 1) / 0.988 + 1
```

Relative humidity model: 
```{r}
# Standard deviation transformation:

# Effect: 
exp(0.314) 

#CI_low: 
exp(0.241)

#CI_high:
exp(0.386)

# sd & unit transformation: 

(exp(0.314) - 1) / 8.46 + 1

```
Relative humidity model 24 h (Version 2): 
```{r}
# Standard deviation transformation:

# Effect: 
exp(0.314) 

#CI_low: 
exp(0.242)

#CI_high:
exp(0.385)

# sd & unit transformation: 

(exp(0.314) - 1) / 8.46 + 1

```

Relative humidity model 48 h (Version 2):
```{r}
# Standard deviation transformation:

# Effect: 
exp(0.308) 

#CI_low: 
exp(0.235)

#CI_high:
exp(0.382)

# sd & unit transformation: 

(exp(0.308) - 1) / 8.46 + 1
```

Moon model: 
```{r}
# Standard deviation transformation:

# Effect: 
exp(-0.128) 

#CI_low: 
exp(-0.194)

#CI_high:
exp(-0.063)

# sd & unit transformation: 

(exp(-0.128) - 1) / 0.355 + 1

```
Night model: 
```{r}
# Standard deviation transformation:

# Effect: 
exp(0.018) 

#CI_low: 
exp(-0.053)

#CI_high:
exp(0.095)

# sd & unit transformation: 
(exp(0.018) - 1) / 0.457 + 1

```
Night model 24h (Version 2): 
```{r}
# Standard deviation transformation:

# Effect: 
exp(0.001) 

#CI_low: 
exp(-0.035)

#CI_high:
exp(0.038)

# sd & unit transformation: 
(exp(0.001) - 1) / 0.457 + 1

```
Night model 48h (Version 2): 
```{r}
# Standard deviation transformation:

# Effect: 
exp(-0.006) 

#CI_low: 
exp(-0.042)

#CI_high:
exp(0.030)

# sd & unit transformation: 
(exp(0.006) - 1) / 0.457 + 1
```

Previous rain 48 h
```{r}
# Standard deviation transformation:

# Effect: 
exp(-0.002) 

#CI_low: 
exp(-0.038)

#CI_high:
exp(0.035)

# sd & unit transformation: 
(exp(-0.002) - 1) / 2.126 + 1

```
Previous rain 48 h (Version 2): 
```{r}
# Standard deviation transformation:

# Effect: 
exp(0.007) 

#CI_low: 
exp(-0.031)

#CI_high:
exp(0.045)

# sd & unit transformation: 
(exp(0.007) - 1) / 0.317 + 1
```

Previous rain 24 h
```{r}
# Standard deviation transformation:

# Effect: 
exp(0.082) 

#CI_low: 
exp(0.045)

#CI_high:
exp(0.121)

# sd & unit transformation: 
(exp(0.082) - 1) / 2.12 + 1
```
Previous rain 24 h rain (Version 2: 
```{r}
# Standard deviation transformation:

# Effect: 
exp(0.090) 

#CI_low: 
exp(0.051)

#CI_high:
exp(0.129)

# sd & unit transformation: 
(exp(0.090) - 1) / 3.16 + 1
```









Moon illumination 
```{r}
# Standard deviation transformation:

# Effect: 
exp(0.128) 

#CI_low: 
exp(0.603)

#CI_high:
exp(0.704)

# sd & unit transformation: 
(exp(0.603) - 1) / 2.12 + 1
```

Previous temp: 
```{r}
# Standard deviation transformation:

# Effect: 
exp(0.016) #-1.016

#CI_low: 
exp(0.083) #-1.086

#CI_high:
exp(0.051)

```





