---
title: <center><font size="6"><b>Statistical analysis</b></font></center>
subtitle: <center><font size="4"><b> Agalychnis lemur</b></font></center>
author: Marcelo Araya-Salas, PhD</a> &  Fabiola Chirino</font></center>
date: <center>`r format(Sys.Date(), "%d-%m-%Y")`</center>
output:
  html_document:
    toc: yes
    toc_depth: 3
    code_folding: hide
    df_print: kable
    toc_float:
      collapsed: no
      smooth_scroll: no
  pdf_document:
    toc: yes
    toc_depth: '3'
fontsize: 12pt
editor_options:
  chunk_output_type: console
---

<!-- this css code centers headings -->
```{css, echo = FALSE}
div#header h1.title, div#header h3.subtitle, div#header h4.author, div#header h4.date {
  text-align: center
}
```

```{r Chunk1 load packages, include=FALSE, message=FALSE, warnings=FALSE, echo=FALSE, eval=TRUE}

# Unload packages
out <- sapply(paste('package:', names(sessionInfo()$otherPkgs), sep = ""), function(x) try(detach(x, unload = FALSE, character.only = TRUE), silent = T))
  
pkgs <- c("remotes", "viridis", "brms", "ggdist", "cowplot", "posterior", "readxl", "HDInterval", "kableExtra", "knitr", "ggdist", "ggplot2", "lunar", "cowplot", "maRce10/brmsish", "warbleR", "ohun")
  
out <- lapply(pkgs, function(y) {
  
  # get pakage name
  pkg <- strsplit(y, "/")[[1]]
  pkg <- pkg[length(pkg)]
  
  # check if installed, if not then install 
  if (!pkg %in% installed.packages()[,"Package"])  {

      if (grepl("/", y))  remotes::install_github(y, force = TRUE) else
    install.packages(y) 
    }

  # load package
  a <- try(require(pkg, character.only = T), silent = T)

  if (!a) remove.packages(pkg)
})

  
options("digits" = 6, "digits.secs" = 5, knitr.table.format = "html") 

# set evaluation false
opts_chunk$set(fig.width = 10, fig.height = 6, warning = FALSE, message = FALSE, tidy = TRUE)

# set working directory as project directory or one directory above,
opts_knit$set(root.dir = "..")

```

```{r functions, eval = TRUE}

source("~/Dropbox/R_package_testing/brmsish/R/extended_summary.R")
source("~/Dropbox/R_package_testing/brmsish/R/helpers.R")

label_spectro_temp <-
  function(wave,
           reference = NULL,
           detection = NULL,
           envelope = FALSE,
           threshold = NULL,
           smooth = 5,
           collevels = seq(-100, 0, 5),
           palette = viridis::viridis,
           template.correlation = NULL,
           line.x.position = 2,
           hop.size = NULL,
           col.line = NULL,
           ...) {
    # adjust wl based on hope.size
    if (!is.null(hop.size))
      wl <- round(wave@samp.rate * hop.size / 1000, 0)

    # reset graphic device on exit
    oldpar <- par(no.readonly = TRUE)
    on.exit(par(oldpar))

    if (envelope | !is.null(template.correlation))
      par(mfrow = c(2, 1), mar = c(0,  4,  1,  1))
    else
      par(mar = c(4, 4, 1, 1))

    # plot spectrogram
    seewave::spectro(
      wave = wave,
      grid = FALSE,
      scale = FALSE,
      palette = palette,
      collevels = collevels,
      axisX = if (envelope |
                  !is.null(template.correlation))
        FALSE
      else
        TRUE,
      ...
    )

    # plot detection
    if (!is.null(reference))
      for (i in seq_len(nrow(reference)))
        lines(
          x = (reference[i, c("start", "end")]),
          y = rep(line.x.position, 2),
          col = col.line[1],
          lwd = 7,
          lend = 2
        )

    # plot detection
    if (!is.null(detection))
      for (i in seq_len(nrow(detection)))
        lines(
          x = (detection[i, c("start", "end")]),
          y = rep(line.x.position - 0.3, 2),
          col = col.line[2],
          lwd = 7,
          lend = 2
        )

    usr <- par("usr")

    # add legend
    if (!is.null(detection) & !is.null(reference))
      legend(
        x = usr[2] * 0.98,
        y = usr[4] * 0.98,
        col = col.line[1:2],
        legend = c("Referencia", "Detección"),
        text.width = 0.3,
        lwd = 4,
        bg = "#FFFFFFE6",
        xjust = 1,
        yjust = 1
      )

    if (is.null(detection) & !is.null(reference))
      legend(
        x = usr[2] * 0.98,
        y = usr[4] * 0.98,
        col = col.line[1],
         text.width = 0.3,
        legend = c("Referencia"),
        lwd = 4,
        bg = "#FFFFFFE6",
        xjust = 1,
        yjust = 1
      )

    if (is.null(reference) & !is.null(detection))
      legend(
        x = usr[2] * 0.98,
        y = usr[4] * 0.98,
        col = col.line[2],
        legend = c("Detección"),
        lwd = 4,
         text.width = 0.3,
        bg = "#FFFFFFE6",
        xjust = 1,
        yjust = 1
      )

    if (envelope) {
      # set graphic device for envelope
      par(mar = c(4,  4,  0.3,  1))

      if (!is.null(smooth))
        smooth <- round(wave@samp.rate * smooth  / 1000, 0)

      # plot envelope
      seewave::env(wave, colwave = "#07889B", ssmooth = smooth)

      # add threshold line
      if (!is.null(threshold))
        abline(h = par("usr")[4] * threshold / 100,
               col = col.line[3],
               lwd = 3)
    } else
      if (!is.null(template.correlation)) {
        # set graphic device for correlations
        par(mar = c(4,  4,  0.3,  1))

        plot(
          x = seq(
            template.correlation$template.duration / 2,
            duration(wave) - template.correlation$template.duration / 2,
            length.out = length(template.correlation$correlation.scores)
          ),
          y = template.correlation$correlation.scores,
          type = "l",
          xlab = "Tiempo (s)",
          ylab = "Correlación",
          col = "#07889B",
          lwd = 1.6,
          xaxs = "i",
          xlim  = c(0, duration(wave))
        )

        # add threshold line
        if (!is.null(threshold))
          abline(h = threshold, col = col.line[3], lwd = 3)

      }
  }
```

```{r, eval = FALSE}

clim_dat_2020 <- read_excel("./data/datos_metereologicos/Estacion_met/Dat_met_estacion_2022-01-11.xlsx",sheet = "2020")

clim_dat_2019 <- read_excel("./data/datos_metereologicos/Estacion_met/Dat_met_estacion_2022-01-11.xlsx",sheet = "2019")

clim_dat <- rbind(clim_dat_2019, clim_dat_2020)

clim_dat <- clim_dat[, c("filename", "Año", "Mes", "Día", "Hora", "Temp (°C)", "Humedad Relat.", "Precipitación")]

names(clim_dat) <- c("filename", "year", "month", "day", "hour", "temp", "HR", "rain")

clim_dat <- aggregate(cbind(rain, temp, HR) ~ filename + year + month + day + hour, clim_dat, mean)

clim_dat$year <- clim_dat$year + 2000

clim_dat$date <- as.Date(paste(clim_dat$year, clim_dat$month, clim_dat$day, sep = "-"))


clim_dat$date_hour <- paste(gsub("-", "", clim_dat$date), clim_dat$hour, sep = "-")

call_dat <- read.csv("./data/processed/call_rate_per_date_time_and_site.csv")

# remove 5 pm data and keep only form Sukia
call_dat <- call_dat[call_dat$hour != 17, ]
call_dat <- call_dat[call_dat$site != "LAGCHIMU", ]


call_dat$date_hour <- paste(sapply(as.character(call_dat$site_date_hour), function(x) strsplit(x, "_")[[1]][2]), call_dat$hour, sep = "-") 


call_dat$temp <- sapply(1:nrow(call_dat), function(x){
  y <- clim_dat$temp[clim_dat$date_hour == call_dat$date_hour[x]]
  
  if (length(y) < 1) y <- NA
  
  return(y)
  }
  )


call_dat$HR <- sapply(1:nrow(call_dat), function(x){
  y <- clim_dat$HR[clim_dat$date_hour == call_dat$date_hour[x]]
  
  if (length(y) < 1) y <- NA
  
  return(y)
  }
  )


call_dat$rain <- sapply(1:nrow(call_dat), function(x){
  y <- clim_dat$rain[clim_dat$date_hour == call_dat$date_hour[x]]
  
  if (length(y) < 1) y <- NA
  
  return(y)
  }
  )

# proportion of acoustic data with climatic data
sum(call_dat$date_hour %in% clim_dat$date_hour) / nrow(call_dat)
sum(!is.na(call_dat$temp)) / nrow(call_dat)


call_dat <- call_dat[!is.na(call_dat$temp), ]

call_dat$day <- as.numeric(substr(sapply(as.character(call_dat$site_date_hour), function(x) strsplit(x, "_")[[1]][2]), 7, 8))

call_dat$date <- as.Date(paste(call_dat$year, call_dat$month, call_dat$day, sep = "-"))

call_dat$moon.date <- ifelse(call_dat$hour < 12, as.Date(call_dat$date - 1), as.Date(call_dat$date))

call_dat$moon.date <- as.Date(call_dat$moon.date, origin = "1970-01-02")

## add moon
call_dat$moonlight <- lunar.illumination(call_dat$moon.date, shift = -6)


call_dat$date_hour_min <- strptime(paste(paste(call_dat$year, call_dat$month, call_dat$day, sep = "-"), paste(call_dat$hour, "00", sep = ":")), format="%Y-%m-%d  %H:%M")

call_dat$hour_diff <- as.numeric(call_dat$date_hour_min - min(call_dat$date_hour_min)) / 3600


call_dat$rain_24 <- sapply(1:nrow(call_dat), function(x) sum(clim_dat$rain[strptime(clim_dat$date, format="%Y-%m-%d") == (strptime(call_dat$date[x], format="%Y-%m-%d") - 60 * 60 * 24)]))

call_dat$rain_48 <- sapply(1:nrow(call_dat), function(x) sum(clim_dat$rain[strptime(clim_dat$date, format="%Y-%m-%d") == (strptime(call_dat$date[x], format="%Y-%m-%d") - 60 * 60 * 48)]))

clim_dat$date_hour_min <- strptime(paste(paste(clim_dat$year, clim_dat$month, clim_dat$day, sep = "-"), paste(clim_dat$hour, "00", sep = ":")), format="%Y-%m-%d %H:%M")

clim_dat$hour_diff <- as.numeric(clim_dat$date_hour_min - min(call_dat$date_hour_min)) / 3600

call_dat$prev_temp <- sapply(1:nrow(call_dat), function(x){
  
  # if(call_dat$hour_diff[x] < 48) 
  #   pt <- NA else
      pt <- mean(clim_dat$temp[clim_dat$hour_diff %in% (call_dat$hour_diff[x] - 48):(call_dat$hour_diff[x] - 24)])
    
    return(pt)
})

write.csv(call_dat, "./data/processed/acoustic_and_climatic_data_by_hour.csv")

```

&nbsp; 

<!-- skyblue box -->

<div class="alert alert-info">

# Purpose

- Evaluate effect of enviromental factors on vocal activity of *A. lemur*

</div>

&nbsp;

## Prepare data

```{r by minute, eval = FALSE, include = FALSE}

clim_dat_2020 <- read_excel("./data/datos_metereologicos/Estacion_met/Dat_met_estacion_2022-01-11.xlsx",sheet = "2020")

clim_dat_2019 <- read_excel("./data/datos_metereologicos/Estacion_met/Dat_met_estacion_2022-01-11.xlsx",sheet = "2019")

clim_dat <- rbind(clim_dat_2019, clim_dat_2020)

clim_dat <- clim_dat[, c("filename", "Año", "Mes", "Día", "Hora", "Min1", "Temp (°C)", "Humedad Relat.", "Precipitación")]

names(clim_dat) <- c("filename", "year", "month", "day", "hour","min", "temp", "HR", "rain")

clim_dat$year <- clim_dat$year + 2000

clim_dat$date <- as.Date(paste(clim_dat$year, clim_dat$month, clim_dat$day, sep = "-"))

clim_dat$date_hour <- paste(gsub("-", "", clim_dat$date), clim_dat$hour, sep = "-")

call_dat <- read.csv("./data/processed/acoustic_and_climatic_data_by_hour.csv")

sub_clim <- clim_dat[clim_dat$date_hour %in% unique(call_dat$date_hour), ]

sub_clim$n_call <- sapply(sub_clim$date_hour, function(x)  call_dat$n_call[call_dat$date_hour == x])

sub_clim$rec_time <- sapply(sub_clim$date_hour, function(x)  call_dat$rec_time[call_dat$date_hour == x])

sub_clim$call_rate <- sapply(sub_clim$date_hour, function(x)  call_dat$call_rate[call_dat$date_hour == x])

sub_clim$rain_24 <- sapply(sub_clim$date_hour, function(x)  call_dat$rain_24[call_dat$date_hour == x])

sub_clim$rain_48 <- sapply(sub_clim$date_hour, function(x)  call_dat$rain_48[call_dat$date_hour == x])

sub_clim$moonlight <- sapply(sub_clim$date_hour, function(x)  call_dat$moonlight[call_dat$date_hour == x])

sub_clim$date_hour_min <- strptime(paste(paste(sub_clim$year, sub_clim$month, sub_clim$day, sep = "-"), paste(sub_clim$hour, sub_clim$min, sep = ":")), format="%Y-%m-%d  %H:%M")

sub_clim$min_diff <- as.numeric(sub_clim$date_hour_min - min(sub_clim$date_hour_min)) / 60

write.csv(sub_clim, "./data/processed/acoustic_and_climatic_data_by_min.csv", row.names = FALSE)

```

## Descriptive stats
```{r}

call_dat_site <- read.csv("./data/processed/call_rate_per_date_time_and_site.csv")

```

- Total number of recordings: `r nrow(call_dat_site)`
- Total recordings per site:

```{r, results='asis'}

agg_recs <- aggregate(rec_time ~ site, data = call_dat_site, length)
names(agg_recs)[1:2] <- c("site", "rec_count")

# print table as kable  
kb <-kable(agg_recs, row.names = TRUE, digits = 3) 
  
kb <- kable_styling(kb, bootstrap_options = c("striped", "hover", "condensed", "responsive"))
  
print(kb)

```

- Total recording time: `r round(sum(call_dat_site$rec_time) / 60)`

- Total recording time per site:

```{r, results='asis'}

agg_recs <- aggregate(rec_time ~ site, data = call_dat_site, sum)
names(agg_recs)[1:2] <- c("site", "recording_time")

agg_recs$recording_time <- round((agg_recs$recording_time) / 60)

# print table as kable  
kb <-kable(agg_recs, row.names = TRUE, digits = 3) 
  
kb <- kable_styling(kb, bootstrap_options = c("striped", "hover", "condensed", "responsive"))
  
print(kb)

```

- Total detections: `r sum(call_dat_site$n_call)`

- Total detections per site:

```{r, results='asis'}

agg_recs <- aggregate(n_call ~ site, data = call_dat_site, sum)
names(agg_recs)[1:2] <- c("calls", "count")

# print table as kable  
kb <-kable(agg_recs, row.names = TRUE, digits = 3) 
  
kb <- kable_styling(kb, bootstrap_options = c("striped", "hover", "condensed", "responsive"))
  
print(kb)

```

- Call rate: `r mean(call_dat_site$call_rate)`

- Call rate per site:

```{r, results='asis'}

agg_recs <- aggregate(call_rate ~ site, data = call_dat_site, mean)
agg_recs$sd <- aggregate(call_rate ~ site, data = call_dat_site, sd)[,2]

names(agg_recs)[1:3] <- c("site", "call_rate", "sd")

# print table as kable  
kb <-kable(agg_recs, row.names = TRUE, digits = 3) 
  
kb <- kable_styling(kb, bootstrap_options = c("striped", "hover", "condensed", "responsive"))
  
print(kb)

```

```{r, eval = TRUE, results='asis'}

call_rate_hour <- read.csv("./data/processed/acoustic_and_climatic_data_by_hour.csv")


agg <- aggregate(cbind(temp, prev_temp, HR, rain, rain_24, rain_48, moonlight)~1, call_rate_hour, function(x) round(c(mean(x), sd(x), min(x), max(x)), 3))

agg <- as.data.frame(matrix(unlist(agg), ncol = 4, byrow = TRUE, dimnames = list(c("Temperature", "Previous temperature", "Relative humidity", "Night rain", "Rain 24 hours", "Rain 48 hours", "Moonlight"),c("mean", "sd", "min", "max"))))
# print table as kable  
kb <-kable(agg, row.names = TRUE, digits = 3) 
  
kb <- kable_styling(kb, bootstrap_options = c("striped", "hover", "condensed", "responsive"))
  
print(kb)

```

Mean and sd temperature: `r mean(call_rate_hour$temp)`

Mean previous temperature: `r mean(call_rate_hour$prev_temp)`

Mean temperature: `r mean(call_rate_hour$temp)`

Mean cumulative rain per hour: `r mean(call_rate_hour$rain)`

Mean cumulative rain per hour previous 24 hours: `r mean(call_rate_hour$rain_24)`

Mean daily cumulative rain per hour previous 48 hours: `r mean(call_rate_hour$rain_48)`




# Bayesian regression models

Scale variables and set model parameters 
```{r prepare data for models, eval = TRUE}

call_rate_hour <- read.csv("./data/processed/acoustic_and_climatic_data_by_hour.csv")

# make hour a factor
call_rate_hour$hour <- factor(call_rate_hour$hour)

# scale and mean-center
call_rate_hour$sc_temp <- scale(call_rate_hour$temp)
call_rate_hour$sc_HR <- scale(call_rate_hour$HR)
call_rate_hour$sc_rain <- scale(call_rate_hour$rain)
call_rate_hour$sc_rain_24 <- scale(call_rate_hour$rain_24)
call_rate_hour$sc_rain_48 <- scale(call_rate_hour$rain_48)
call_rate_hour$sc_moonlight <- scale(call_rate_hour$moonlight)
call_rate_hour$sc_prev_temp <- scale(call_rate_hour$prev_temp)

priors <- c(prior(normal(0, 4), class = "b"))
chains <- 4
iter <- 10000

```

## Global models

The increase in relative humidity, decrease in temperature, increase in the previous accumulated rain, decrease in the night rain, decrease in the percentage of the moon illuminated cause the activity of A. lemur to increase? 

```{r, eval = FALSE}

fit_glob1 <- brm(n_call | resp_rate(rec_time) ~  sc_temp + sc_HR + sc_moonlight + sc_rain + sc_rain_24 + sc_prev_temp + ar(p = 2, time = hour_diff, gr = hour), data = call_rate_hour, iter = iter, chains = chains, cores = chains, family = negbinomial(), prior = priors, file = "./data/processed/regression_models/global_rain_24", file_refit = "always")


fit_glob2 <- brm(n_call | resp_rate(rec_time) ~  sc_temp + sc_HR + sc_moonlight + sc_rain + sc_rain_48 + sc_prev_temp + ar(p = 2, time = hour_diff, gr = hour), data = call_rate_hour, iter = iter, chains = chains, cores = chains, family = negbinomial(), prior = priors, file = "./data/processed/regression_models/global_rain_48", file_refit = "always")


```

```{r, eval = TRUE, results='asis'}
 
extended_summary(read.file = "./data/processed/regression_models/global_rain_24.rds", n.posterior = 1000)

extended_summary(read.file = "./data/processed/regression_models/global_rain_48.rds", n.posterior = 1000)

```


## Temperature

The increase in temperature at night causes that activity of A. lemur to decrease? 

```{r, eval= FALSE}

fit2.1 <- brm(n_call | resp_rate(rec_time) ~  sc_temp + sc_rain + sc_rain_24 + ar(p = 2, time = hour_diff, gr = hour), data = call_rate_hour, iter = iter, chains = chains, cores = chains, family = negbinomial(), prior = priors, file = "./data/processed/regression_models/temp2.1", file_refit = "always") 

fit2.2 <- brm(n_call | resp_rate(rec_time) ~  sc_temp + sc_rain + sc_rain_48 + ar(p = 2, time = hour_diff, gr = hour), data = call_rate_hour, iter = iter, chains = chains, cores = chains, family = negbinomial(), prior = priors, file = "./data/processed/regression_models/temp2.2", file_refit = "always")

fit2.3 <- brm(n_call | resp_rate(rec_time) ~ sc_prev_temp + ar(p = 2, time = hour_diff, gr = hour), data = call_rate_hour, iter = iter, chains = chains, cores = chains, family = negbinomial(), prior = priors, file = "./data/processed/regression_models/temp2.3", file_refit = "always")

```

```{r, eval = TRUE, results='asis'}

extended_summary(read.file = "./data/processed/regression_models/temp2.1.rds", n.posterior = 1000)

extended_summary(read.file = "./data/processed/regression_models/temp2.2.rds", n.posterior = 1000)

extended_summary(read.file = "./data/processed/regression_models/temp2.3.rds", n.posterior = 1000)
```

##  Relative humidity

Does an increase in relative humidity cause the activity of A. lemur to increase? 

```{r, eval= FALSE}

fit.3.1 <- brm(n_call | resp_rate(rec_time) ~  sc_temp + sc_HR + sc_rain + sc_rain_24  + ar(p = 2, time = hour_diff, gr = hour), data = call_rate_hour, iter = iter, chains = chains, cores = chains, family = negbinomial(), prior = priors, file = "./data/processed/regression_models/RH3.1", file_refit = "always")

fit.3.2 <- brm(n_call | resp_rate(rec_time) ~  sc_temp + sc_HR + sc_rain + sc_rain_48  + ar(p = 2, time = hour_diff, gr = hour), data = call_rate_hour, iter = iter, chains = chains, cores = chains, family = negbinomial(), prior = priors, file = "./data/processed/regression_models/RH3.2", file_refit = "always") # Este tengo que doblemente confirmarlo
```

```{r, eval = TRUE, results='asis'}

extended_summary(read.file = "./data/processed/regression_models/RH3.1.rds", n.posterior = 1000)

extended_summary(read.file = "./data/processed/regression_models/RH3.2.rds", n.posterior = 1000)

```

## Moon

Decreasing the percentage of the moon illuminated causes an increase in A. lemur activity?

```{r, eval= FALSE}

fit.4 <- brm(n_call | resp_rate(rec_time) ~  sc_moonlight + ar(p = 2, time = hour_diff, gr = hour), data = call_rate_hour, iter = iter, chains = chains, cores = chains, family = negbinomial(), prior = priors, file = "./data/processed/regression_models/moon4", file_refit = "always")

```

```{r, eval = TRUE, results='asis'}

extended_summary(read.file = "./data/processed/regression_models/moon4.rds", n.posterior = 1000)

```

## Night rain

If the night rain decreases can affect the A. lemur activity to increase?

```{r, eval= FALSE}

fit.5.1 <- brm(n_call | resp_rate(rec_time) ~  sc_prev_temp + sc_rain + sc_rain_48  + ar(p = 2, time = hour_diff, gr = hour), data = call_rate_hour, iter = iter, chains = chains, cores = chains, family = negbinomial(), prior = priors, file = "./data/processed/regression_models/night_rain5.1", file_refit = "always")

fit.5.2 <- brm(n_call | resp_rate(rec_time) ~  sc_prev_temp + sc_rain + sc_rain_24  + ar(p = 2, time = hour_diff, gr = hour), data = call_rate_hour, iter = iter, chains = chains, cores = chains, family = negbinomial(), prior = priors, file = "./data/processed/regression_models/night_rain5.2", file_refit = "always")


```

```{r, eval = TRUE, results='asis'}

extended_summary(read.file = "./data/processed/regression_models/night_rain5.1.rds", n.posterior = 1000)

extended_summary(read.file = "./data/processed/regression_models/night_rain5.2.rds", n.posterior = 1000)

```

## Previous Rain

Does an increase in the accumulated previous rain causes that A. lemur activity to increase?

```{r, eval= FALSE}

fit.6.1 <- brm(n_call | resp_rate(rec_time) ~ sc_temp +  sc_rain  + sc_rain_48 + ar(p = 2, time = hour_diff, gr = hour), data = call_rate_hour, iter = iter, chains = chains, cores = chains, family = negbinomial(), prior = priors, file = "./data/processed/regression_models/previous_rain6.1", file_refit = "always")

fit.6.2 <- brm(n_call | resp_rate(rec_time) ~ sc_temp +  sc_rain + sc_rain_24  + ar(p = 2, time = hour_diff, gr = hour), data = call_rate_hour, iter = iter, chains = chains, cores = chains, family = negbinomial(), prior = priors, file = "./data/processed/regression_models/previous_rain6.2", file_refit = "always")

fit.6.3 <- brm(n_call | resp_rate(rec_time) ~ sc_temp + sc_HR + sc_rain_24 + ar(p = 2, time = hour_diff, gr = hour), data = call_rate_hour, iter = iter, chains = chains, cores = chains, family = negbinomial(), prior = priors, file = "./data/processed/regression_models/night_rain6.3", file_refit = "always")
```

```{r, eval = TRUE, results='asis'}

extended_summary(read.file = "./data/processed/regression_models/previous_rain6.1.rds", n.posterior = 1000)

extended_summary(read.file = "./data/processed/regression_models/previous_rain6.2.rds", n.posterior = 1000)

extended_summary(read.file = "./data/processed/regression_models/night_rain6.3.rds", n.posterior = 1000)

```

# Causal model combined effect size plot

Takes the effect sizes (and posteriors) from the right causal models
```{r}

coef_table <- read.csv("./data/processed/summary_causal_model_table_13-01-23.csv", sep = "\t")

coef_table$variable <- coef_table$Label
coef_table$value <- coef_table$Estimate
coef_table$significance <- ifelse(coef_table$CI_low * coef_table$CI_high > 0, "sig", "no.sig")

posteriors_l <- lapply(1:nrow(coef_table), function(x){

  # print(x)
  X <- readRDS(coef_table$Model[x])
  xdrws <- brms::as_draws(X)
  post <- xdrws$`1`[[paste0("b_", coef_table$Variable[x])]] 
  out <- data.frame(variable = coef_table$Label[x], value = post, significance = coef_table$significance[x])
  return(out)
})

posteriors <- do.call(rbind, posteriors_l)

coef_table$variable <- factor(coef_table$Label, levels = sort(unique(coef_table$Label), FALSE))
posteriors$variable <- factor(posteriors$variable, levels = sort(unique(posteriors$variable), TRUE))


fill_values <- c("#FFB9DF", "#FF7598")
fill_values <- adjustcolor(fill_values, alpha.f = 0.5)

# creat plots
    # gg_dists <-
      ggplot2::ggplot(data = posteriors, ggplot2::aes(y = variable, x = value, fill = significance)) +
      ggplot2::geom_vline(xintercept = 0,
                          col = "black",
                          lty = 2) +
      ggdist::stat_halfeye(
        ggplot2::aes(x = value),
        .width = c(.95),
        normalize = "panels",
        color = "transparent"
      ) +
      ggplot2::scale_fill_manual(values = fill_values, guide = 'none') +
      ggplot2::geom_point(data = coef_table) +
      ggplot2::geom_errorbar(data = coef_table,
                             ggplot2::aes(xmin = CI_low, xmax = CI_high),
                             width = 0) +
      ggplot2::scale_color_manual(values = coef) +
      ggplot2::facet_wrap(
        ~ variable,
        scales = "free_y",
        ncol = 1
      ) +
      ggplot2::theme_classic() +
      ggplot2::theme(
        axis.ticks.length = ggplot2::unit(0, "pt"),
        plot.margin = ggplot2::margin(0, 0, 0, 0, "pt"),
        legend.position = "none",
        strip.background = ggplot2::element_blank(),
        strip.text = ggplot2::element_blank()
      ) +
      ggplot2::labs(x = "Effect size", y = "Parameter") #+

      # ggplot2::xlim(range(c(posteriors_by_chain$value, 0)) * plot.area.prop)

      ggsave(filename =
               "./figures/summary_effect_sizes_pooled_from_multiple_causal_models.jpeg")
```


```{r, subset for presentation, eval = FALSE}

coef_table <- read.csv("./data/processed/summary_causal_model_table_13-01-23.csv", sep = "\t")

coef_table <- coef_table[coef_table$Label %in% c("Temperature (day-1 rain)", "Previous mean temperature", "Relative humidity (day-1 rain)", "Moonlight", "Night Rain (day-1 rain)", "Previous day-1 rain", "Previous day-2 rain"), ]

coef_table$variable <- coef_table$Label
coef_table$value <- coef_table$Estimate
coef_table$significance <- ifelse(coef_table$CI_low * coef_table$CI_high > 0, "sig", "no.sig")

posteriors_l <- lapply(1:nrow(coef_table), function(x){

  # print(x)
  X <- readRDS(coef_table$Model[x])
  xdrws <- brms::as_draws(X)
  post <- xdrws$`1`[[paste0("b_", coef_table$Variable[x])]] 
  out <- data.frame(variable = coef_table$Label[x], value = post, significance = coef_table$significance[x])
  return(out)
})

posteriors <- do.call(rbind, posteriors_l)

coef_table$variable <- factor(coef_table$Label, levels = sort(unique(coef_table$Label), FALSE))
posteriors$variable <- factor(posteriors$variable, levels = sort(unique(posteriors$variable), TRUE))


fill_values <- c("#FFB9DF", "#FF7598")
fill_values <- adjustcolor(fill_values, alpha.f = 0.5)

# creat plots
    # gg_dists <-
      ggplot2::ggplot(data = posteriors, ggplot2::aes(y = variable, x = value, fill = significance)) +
      ggplot2::geom_vline(xintercept = 0,
                          col = "black",
                          lty = 2) +
      ggdist::stat_halfeye(
        ggplot2::aes(x = value),
        .width = c(.95),
        normalize = "panels",
        color = "transparent"
      ) +
      ggplot2::scale_fill_manual(values = fill_values, guide = 'none') +
      ggplot2::geom_point(data = coef_table) +
      ggplot2::geom_errorbar(data = coef_table,
                             ggplot2::aes(xmin = CI_low, xmax = CI_high),
                             width = 0) +
      ggplot2::scale_color_manual(values = coef) +
      ggplot2::facet_wrap(
        ~ variable,
        scales = "free_y",
        ncol = 1
      ) +
      ggplot2::theme_classic() +
      ggplot2::theme(
        axis.ticks.length = ggplot2::unit(0, "pt"),
        plot.margin = ggplot2::margin(0, 0, 0, 0, "pt"),
        legend.position = "none",
        strip.background = ggplot2::element_blank(),
        strip.text = ggplot2::element_blank()
      ) +
      ggplot2::labs(x = "Effect size", y = "Parameter") #+

      # ggplot2::xlim(range(c(posteriors_by_chain$value, 0)) * plot.area.prop)

      ggsave(filename =
               "./figures/summary_effect_sizes_pooled_from_multiple_causal_models.jpeg")
```

# Conditional plots

Measured based on the global model

## Rain and temperature
```{r, eval = TRUE}

glob_mod_24 <- readRDS("./data/processed/regression_models/global_rain_24.rds")

conditions <- data.frame(sc_temp = c(`Low temperature` = -1, `Mean temperature` = 0, `High temperature` = 1))

rain_gg <- plot(conditional_effects(glob_mod_24, effects = "sc_rain", 
                         conditions = conditions), plot = FALSE)[[1]] +
  scale_color_viridis_d() + theme_classic() + ylim(c(0, 520)) + ggtitle("Current rain") +  labs(x = "Rain", y = "Call activity (calls/hour)")

rain24_gg <- plot(conditional_effects(glob_mod_24, effects = "sc_rain_24", 
                         conditions = conditions), plot = FALSE)[[1]] +
  scale_color_viridis_d() + theme_classic() + ylim(c(0, 520)) + ggtitle("Previous 24h rain") + labs(x = "Rain", y = "Call activity (calls/hour)") 

glob_mod_48 <- readRDS("./data/processed/regression_models/global_rain_48.rds")

rain48_gg <- plot(conditional_effects(glob_mod_48, effects = "sc_rain_48", 
                         conditions = conditions), plot = FALSE)[[1]] +
  scale_color_viridis_d() + theme_classic() + ylim(c(0, 520)) + ggtitle("Previous 48h rain") + labs(x = "Rain", y = "Call activity (calls/hour)")


cowplot::plot_grid(rain_gg, rain24_gg, rain48_gg, nrow = 3)

```

## Relative humidity and temperature
```{r, eval = TRUE}

conditions <- data.frame(sc_temp = c(`Low temperature` = -1, `Mean temperature` = 0, `High temperature` = 1))

hr_temp_gg <- plot(conditional_effects(glob_mod_24, effects = "sc_HR", 
                         conditions = conditions), plot = FALSE)[[1]] +
  scale_color_viridis_d() + theme_classic() + labs(x = "Relative humidity", y = "Call activity (calls/hour)")

hr_temp_gg

conditions <- data.frame(sc_HR = c(`Low humidity` = -1, `Mean humidity` = 0, `High humidity` = 1))

temp_hr_gg <- plot(conditional_effects(glob_mod_24, effects = "sc_temp", 
                         conditions = conditions), plot = FALSE)[[1]] +
  scale_color_viridis_d() + theme_classic() + labs(x = "Temperature", y = "Call activity (calls/hour)")

temp_hr_gg

```

## Relative humidity and rain
```{r, eval = TRUE}

conditions <- data.frame(sc_rain = c(`Low rain` = -1, `Mean rain` = 0, `High rain` = 1))


rain_gg <- plot(conditional_effects(glob_mod_24, effects = "sc_HR", 
                         conditions = conditions), plot = FALSE)[[1]] +
  scale_color_viridis_d() + theme_classic() + ylim(c(0, 75)) + ggtitle("Current rain")

rain24_gg <- plot(conditional_effects(glob_mod_24, effects = "sc_HR", 
                         conditions = conditions), plot = FALSE)[[1]] +
  scale_color_viridis_d() + theme_classic() + ylim(c(0, 75)) + ggtitle("Previous 24h rain") 

rain48_gg <- plot(conditional_effects(glob_mod_48, effects = "sc_HR", 
                         conditions = conditions), plot = FALSE)[[1]] +
  scale_color_viridis_d() + theme_classic() + ylim(c(0, 75)) + ggtitle("Previous 48h rain") 

cowplot::plot_grid(rain_gg, rain24_gg, rain48_gg, nrow = 3)


conditions <- data.frame(sc_HR = c(`Low humidity` = -1, `Mean humidity` = 0, `High humidity` = 1))

rain_hr_gg <- plot(conditional_effects(glob_mod_24, effects = "sc_rain", 
                         conditions = conditions), plot = FALSE)[[1]] +
  scale_color_viridis_d() + theme_classic() + ggtitle("Current rain") + labs(x = "Rain", y = "Call activity (calls/hour)")

rain24_hr_gg <- plot(conditional_effects(glob_mod_24, effects = "sc_rain_24", 
                         conditions = conditions), plot = FALSE)[[1]] +
  scale_color_viridis_d() + theme_classic() + ggtitle("Previous 24h rain") + labs(x = "Rain previous", y = "Call activity (calls/hour)")

rain48_hr_gg <- plot(conditional_effects(glob_mod_48, effects = "sc_rain_48", 
                         conditions = conditions), plot = FALSE)[[1]] +
  scale_color_viridis_d() + theme_classic() + ggtitle("Previous 48h rain") + labs(x = "Rain previous", y = "Call activity (calls/hour)")

cowplot::plot_grid(rain_hr_gg, rain24_hr_gg, rain48_hr_gg, nrow = 3)

```

## Moonlight and temperature

```{r, eval = TRUE}

conditions <- data.frame(sc_temp = c(`Low temperature` = -1, `Mean temperature` = 0, `High temperature` = 1))

moon_temp_gg <- plot(conditional_effects(glob_mod_24, effects = "sc_moonlight", 
                         conditions = conditions), plot = FALSE)[[1]] +
  scale_color_viridis_d() + theme_classic() + labs(x = "Moonlight", y = "Call activity (calls/hour)")

moon_temp_gg

conditions <- data.frame(sc_moonlight = c(`Low moonlight` = -1, `Mean moonlight` = 0, `High moonlight` = 1))

temp_moon_gg <- plot(conditional_effects(glob_mod_24, effects = "sc_temp", 
                         conditions = conditions), plot = FALSE)[[1]] +
  scale_color_viridis_d() + theme_classic() + labs(x = "Temperature", y = "Call activity (calls/hour)")

temp_moon_gg

```


```{r, eval = FALSE}
# st <- data.frame(sound.files = "FIGURA_LEMUR.wav", selec = 1, start = 0.1, end = 0.4)

# tweak_spectro(st, length.out = 20, ovlp = 99, wl = c(100, 1000), pal = c("reverse.gray.colors.2", "viridis", "reverse.terrain.colors"), path = "./figures/Figura_canto_lemur", flim = c(1, 4), ncol = 10, nrow = 6, width = 15, collev.min = c(-110))

wav <- readWave("./figures/Figura_canto_lemur/FIGURA_LEMUR.wav")

graphics.off()
par(mfrow = c(2, 1), mar = c(0, 4, 4, 1))

warbleR:::spectro_wrblr_int2(wav, flim = c(1, 4), wl = 700, grid = F, collevels = seq(-125, 0 ,1), ovlp = 0, palette = viridis, axisX = FALSE)

peaks <- c(0.320165, 0.9, 1.7, 2.6, 3.5, 4.2, 4.5) # + seq(-0.1, 0.2, length.out = 7)
valleys <- seq(0, duration(wav) + 0.3, length.out = 100)

cor_dat <- data.frame(time = c(peaks, valleys), cor = c(rep(0.7, length(peaks)), rep(0.1, length(valleys))))

cor_dat$cor <- cor_dat$cor  + rnorm(nrow(cor_dat), sd = 0.03)
cor_dat$cor <- smoothw(cor_dat$cor, wl=2, f = 10)


cor_dat <- cor_dat[order(cor_dat$time),]

par(mar = c(2, 4, 0, 1))

plot(cor_dat$time, cor_dat$cor, type = "l", xaxs="i")



spectro(wav, flim = c(1, 4), wl = 700, grid = F, collevels = seq(-125, 0 ,1), ovlp = 99, palette = viridis, axisX = FALSE, tlim = c(1.53, 1.63), scale = FALSE, flab = "Frecuencia (kHz)", tlab = "Tiempo (s)")

```


```{r, eval = FALSE}

template <- data.frame(sound.files = "FIGURA_LEMUR.wav", selec = 1, start = 1.55, end = 1.61, bottom.freq = 2, top.freq = 3)


# get correlations
correlations <-
  template_correlator(templates = template,
                      files = "FIGURA_LEMUR.wav",
                      path = "./figures/Figura_canto_lemur/")


thresh <- 0.7
# run detection
detection <-
  template_detector(template.correlations = correlations, threshold = thresh)

reference <-
  template_detector(template.correlations = correlations, threshold = 0.55)
detection


# plot spectrogram
label_spectro_temp(
  wave = wav,
  reference = reference,
  detection = detection,
  template.correlation = correlations[[1]],
  flim = c(1, 4),
  threshold = thresh,
  hop.size = 10, ovlp = 50, collevels = seq(-125, 0 ,1), 
  col.line = c("#AFBF35","#F20505", "#F2622E"),
   flab = "Frecuencia (kHz)", tlab = "Tiempo (s)")

#012623 
#034941 
#AFBF35 
#F2622E 
#F20505



```

---


<!-- light green box -->

<div class="alert alert-success">

&nbsp; 

# Takeaways

- 

</div>

&nbsp;

<div class="alert alert-info">

# Sum up results

-  

</div>


# Next steps

- 

&nbsp; 

---

<font size="4">Session information</font>

```{r session info, echo=F, eval = TRUE}

sessionInfo()

```
