---
title: "Template validation"
author: "Beatriz Willink y Fabiola Chirino"
date: "10/9/2020"
output: 
  html_document:
    theme: united
    highlight: tango
    number_sections: true
    toc: true
    toc_float: true
    
---
```{r setup, include=FALSE,eval=FALSE}
knitr::opts_knit$set(root.dir = "..", eval = TRUE)
```


Load required packages 
```{r, message=FALSE}
require(monitoR)
require(ggplot2)
require(wesanderson)
```


```{r message=FALSE}
timeAlign2 <- function (x, what = "detections", tol = 1) 
{
  tol <- 0.5 * tol
  if (class(x) == "detectionList" && tolower(what) %in% c("detections", 
                                                          "d", "det")) {
    cat("Getting detections\n")
    all.results <- getDetections(x)
  }
  else if (class(x) == "detectionList" && tolower(what) %in% 
           c("peaks", "p", "pks")) {
    cat("Getting peaks\n")
    all.results <- getPeaks(x)
  }
  else if (class(x) == "data.frame") {
    all.results <- x
  }
  else if (class(x) == "list") {
    check.class <- unlist(lapply(as.list(x), function(file) class(file)))
    if (all(check.class == "character")) {
      all.results <- lapply(x, function(file) {
        chars <- nchar(file)
        ext <- tolower(gsub(".*\\.", "", file))
        if (all(ext == "csv")) {
          all.results <- lapply(x, function(data) read.csv(data, 
                                                           stringsAsFactors = FALSE))
        }
        else stop("x must be a csv file.")
        return(all.results)
      })
    }
    else if (all(check.class == "data.frame")) 
      all.results <- rbindf(x)
  }
  else stop("x is unfamiliar; must be detectionList, data frame, or a list of either data frames or file paths to csv files.")
  colNm <- names(all.results)
  all.results <- all.results[, colNm]
  names.res <- names(all.results)
  if (all(c("fldTime", "fldTemplateName", "fldScore") %in% 
          names.res)) {
    time.fld <- "fldTime"
    temp.fld <- "fldTemplateName"
    scor.fld <- "fldScore"
    onAmp.fld <- "fldOnAmp"
    offAmp.fld <- "fldOffAmp"
  }
  else if (all(c("time", "template", "score") %in% names.res)) {
    time.fld <- "time"
    temp.fld <- "template"
    scor.fld <- "score"
    onAmp.fld <- "on.amp"
    offAmp.fld <- "off.amp"
  }
  else stop("Unrecognized column names.  x must be from findPeaks or from an acoustics database.")
  if (any(c("fldOnAmp", "fldOffAmp", "on.amp", "off.amp") %in% 
          names.res)) {
    report.amp <- TRUE
  }
  else report.amp <- FALSE
  all.results <- all.results[order(all.results[, time.fld]), 
                             ]
  row.names(all.results) <- 1:nrow(all.results)
  nRows <- nrow(all.results)
  earlytimes <- all.results[c(1:(nRows - 1)), time.fld]
  latetimes <- all.results[c(2:nRows), time.fld]
  same1below <- c(earlytimes + tol >= latetimes & earlytimes + 
                    tol <= latetimes + tol, FALSE)
  same1above <- c(FALSE, earlytimes + tol >= latetimes & earlytimes + 
                    tol <= latetimes + tol)
  group <- same1above * same1below
  nTimes <- nScore <- nOnAmp <- nOffAmp <- nTemplateName <- rep(NA, 
                                                                nRows)
  groupID <- rep(NA, nRows)
  gid <- 1
  if (nRows > 2) {
    for (i in 2:(nRows - 1)) {
      if (group[i]) {
        groupID[(i - 1):(i + 1)] <- gid
      }
      else if (!group[i] & group[i - 1])
        gid <- gid + 1
    }
  } else
  {
    for (i in 2:2) {
      if (group[i]) {
        groupID[(i - 1):(i + 1)] <- gid
      }
      else if (!group[i] & group[i - 1])
        gid <- gid + 1
    }
  }
  for (i in na.omit(unique(groupID))) {
    rID <- which(groupID == i)
    num.evt <- length(rID)
    rmaxscor <- rID[which(all.results[rID, scor.fld] == 
                            max(all.results[rID, scor.fld]))]
    if (length(rmaxscor) > 1) 
      rmaxscor <- rmaxscor[1]
    nScore[rID] <- rep(all.results[rmaxscor, scor.fld], 
                       num.evt)
    nTimes[rID] <- rep(all.results[rmaxscor, time.fld], 
                       num.evt)
    if (report.amp) {
      nOnAmp[rID] <- rep(all.results[rmaxscor, onAmp.fld], 
                         num.evt)
      nOffAmp[rID] <- rep(all.results[rmaxscor, offAmp.fld], 
                          num.evt)
    }
    nTemplateName[rID] <- rep(all.results[rmaxscor, temp.fld], 
                              num.evt)
  }
  groupID <- rep(NA, nRows)
  gid <- 0
  for (i in 1:nRows) {
    if (any(same1above[i], same1below[i]) && is.na(nScore[i])) {
      if (same1above[i]) 
        groupID[i] <- gid
      else {
        gid <- gid + 1
        groupID[i] <- gid
      }
    }
  }
  for (i in na.omit(unique(groupID))) {
    rID <- which(groupID == i)
    num.evt <- length(rID)
    rmaxscor <- rID[which(all.results[rID, scor.fld] == 
                            max(all.results[rID, scor.fld]))]
    if (length(rmaxscor) > 1) 
      rmaxscor <- rmaxscor[1]
    nScore[rID] <- rep(all.results[rmaxscor, scor.fld], 
                       num.evt)
    nTimes[rID] <- rep(all.results[rmaxscor, time.fld], 
                       num.evt)
    if (report.amp) {
      nOnAmp[rID] <- rep(all.results[rmaxscor, onAmp.fld], 
                         num.evt)
      nOffAmp[rID] <- rep(all.results[rmaxscor, offAmp.fld], 
                          num.evt)
    }
    nTemplateName[rID] <- rep(all.results[rmaxscor, temp.fld], 
                              num.evt)
  }
  rID <- which(is.na(nScore))
  nScore[rID] <- all.results[rID, scor.fld]
  nTimes[rID] <- all.results[rID, time.fld]
  if (report.amp) {
    nOnAmp[rID] <- all.results[rID, onAmp.fld]
    nOffAmp[rID] <- all.results[rID, offAmp.fld]
  }
  nTemplateName[rID] <- all.results[rID, temp.fld]
  all.results[time.fld] <- nTimes
  all.results[scor.fld] <- nScore
  if (report.amp) {
    all.results[onAmp.fld] <- nOnAmp
    all.results[offAmp.fld] <- nOffAmp
  }
  all.results[temp.fld] <- nTemplateName
  goodRows <- cbind(all.results[time.fld], all.results[scor.fld])
  goodRows <- unique(goodRows)
  goodRows <- as.numeric(row.names(goodRows))
  all.results <- all.results[goodRows, ]
  row.names(all.results) <- 1:nrow(all.results)
  return(all.results)
}






```


```{r, message=FALSE}
# Set .wav player
setWavPlayer("play")
```

## Rename recordings

This helps MonitoR read the recording metadata correctly. The function fileCopyRename changed the dates in the metadata, so we do it manually (for now).

### **Note**

All recordings have been previously copied to the directory "Fixed_names".

This should work with relative paths, but it doesn't... So we are using the absolute path instead
```{r}

getwd()

path.start= "../"
```


```{r, message=FALSE, eval=FALSE}
# fnames <-
#   dir(path = paste0(path.start, "data/Processed/Fixed_Names/"),
#       full.names = T,
#       recursive = T)
# 
# # Agregamos un guion despues del anio y mes
# n.fnames <-
#   gsub(pattern = "(A_[0-9]{4})" ,
#        replacement = "\\1-",
#        x = fnames)
# n.fnames <-
#   gsub(pattern = "(A_[0-9]{4}-[0-9]{2})" ,
#        replacement = "\\1-",
#        x = n.fnames)
# 
# #Agregamos la zona horaria
# n.fnames <-
#   gsub(pattern = "(.wav)" ,
#        replacement = "_UTC-6\\1",
#        x = n.fnames)

#Actualizamos los nombres
#file.rename(fnames, n.fnames)
```

## Clip 5 min surveys

This is needed for old recordings that are 20 min long. We only need the first and final 5 min of each
```{r, eval=TRUE}
# First 5 min
for (i in 1:length(n.fnames)) {
  # read in the first 5 min
  rectemp <-
    readWave(paste0(n.fnames[i]),
             from = 0,
             to = 300,
             units = "seconds")
  
  # change time to end of recording
  f.time <- gsub(
    pattern = "0000",
    replacement = "0500",
    x = paste0(n.fnames[i])
  )
  
  # change directory for trimmed recordings
  n.name <-
    gsub(pattern = "Fixed_Names/",
         replacement = "Five_Min_Recs/",
         x = f.time)
  
  #write file with new name
  writeWave(object = rectemp, filename = paste0(n.name))
}

# Final 5 min
for (i in 1:length(n.fnames)) {
  # read in the first 5 min
  rectemp <-
    readWave(paste0(n.fnames[i]),
             from = 900,
             to = 1200,
             units = "seconds")
  
  # change time to end of recording
  f.time <- gsub(
    pattern = "0000",
    replacement = "2000",
    x = paste0(n.fnames[i])
  )
  
  # change directory for trimmed recordings
  n.name <-
    gsub(pattern = "Fixed_Names/",
         replacement = "Five_Min_Recs/",
         x = f.time)
  
  #write file with new name
  writeWave(object = rectemp, filename = paste0(n.name))
}

# Create monitoR metadata in case we need them later
# Again we are using absolute paths out of need
fileCopyRename(
  files = dir(
    path = paste0(path.start, "data/Processed/Five_Min_Recs/"),
    full.names = T
  ),
  to = paste0(path.start, "data/Processed/Five_Min_Recs/"),
  csv.dir = paste0(path.start, "data/Processed/Five_Min_Recs/"),
  loc.prefix = "SUKIAP",
  csv.name = "ValidacionMetadata.csv",
  ext = "wav",
  rec.tz = "UTC-6",
  full.survey.names = F,
  metadata.only = T,
  rename = F,
  copy = F
)
```

## Templates

Read in recordings used for templates

```{r, eval= FALSE}




# Template 3: SUKIA_20190906_21000.wav (21:05 2 cantos en 1 mismo segundo)
rec3 <- readWave(paste0(path.start,
                        "data/Raw/SUKIA_20190906_210000.wav"))
rec3<-cutWave(rec3,from=180, to=181)
writeWave(rec3,
          paste0(
            path.start,
            "data/Processed/templates/rec_20190906_21_05pm.wav"
          ))



# Template 5: SUKIA_20191011_21000.wav (21:05, lluvia y actividad lemur)
rec5 <- readWave(paste0(path.start,
                        "data/Raw/SUKIA_20191011_210000.wav"))
rec5 <- cutWave(rec5, from = 7, to = 8)
writeWave(rec5,
          paste0(
            path.start,
            "data/Processed/templates/rec_20191011_21_05pm.wav"
          ))


# Template8: SUKIA_20191102_22000.wav (22:05, poca actividad)
rec8 <- readWave(paste0(path.start,
                        "data/Raw/SUKIA_20191102_220000.wav"))
rec8 <- cutWave(rec8, from = 27, to = 28)
writeWave(rec8,
          paste0(
            path.start,
            "data/Processed/templates/rec_20191102_22_05pm.wav"
          ))

# Template 9: SUKIA_20191101_18000.wav (18:05, poca actividad)
rec9 <- readWave(paste0(path.start,
                        "data/Raw/SUKIA_20191101_180000.wav"))
rec9 <- cutWave(rec9, from = 134, to = 135)
writeWave(rec9,
          paste0(
            path.start,
            "data/Processed/templates/rec_20191101_18_05pm.wav"
          ))

# Template 10: SUKIA_20191204_010000.wav (01:05, actividad de Cruziohyla y Lemur)
rec10 <- readWave(paste0(path.start,
                         "data/Raw/SUKIA_20191204_010000.wav"))
rec10 <- cutWave(rec10, from = 83, to = 84)
writeWave(rec10,
          paste0(
            path.start,
            "data/Processed/templates/rec_20191204_01_05pm.wav"
          ))

# Template 11: SUKIA_20191219_22000.wav (22:05, mucha actividad Lemur)
rec11 <- readWave(paste0(path.start,
                         "data/Raw/SUKIA_20191219_220000.wav"))
rec11 <- cutWave(rec11, from = 10, to = 11)
writeWave(rec11,
          paste0(
            path.start,
            "data/Processed/templates/rec_20191219_22_05pm.wav"
          ))

# Template 12: SUKIA_20191215_040000.wav (04:15, muy poca actividad)
rec12 <- readWave(paste0(path.start,
                         "data/Raw/SUKIA_20191215_040000.wav"))
rec12 <- cutWave(rec12, from = 987, to = 988)
writeWave(rec12,
          paste0(
            path.start,
            "data/Processed/templates/rec_20191215_04_15pm.wav"
          ))



# Template 16: SUKIA_20191126_22000.wav (22:15, actividad #2)
rec16 <- readWave(paste0(path.start,
                         "data/Raw/SUKIA_20191202_200000.wav"))
rec16 <- cutWave(rec16, from = 902, to = 903)
writeWave(rec16,
          paste0(
            path.start,
            "data/Processed/templates/rec_20191202_20_00pm.wav"
          ))

# Template 17: SUKIA_20200102_23000.wav (23:15 mucha mucha actividad)
rec17 <- readWave(paste0(path.start,
                         "data/Raw/SUKIA_20200102_230000.wav"))
rec17 <- cutWave(rec17, from = 911, to = 912)
writeWave(rec17,
          paste0(
            path.start,
            "data/Processed/templates/rec_20200102_23_15pm.wav"
          ))


# Template 19: SUKIA_20200131_030000.wav (03:15)
rec19 <- readWave(paste0(path.start,
                         "data/Raw/SUKIA_20200131_030000.wav"))
rec19 <- cutWave(rec19, from = 977, to = 978)
writeWave(rec19,
          paste0(
            path.start,
            "data/Processed/templates/rec_20200131_03_15pm.wav"
          ))

# Template 21: SUKIA_20200204_040000.wav (04:05)
rec21 <- readWave(paste0(path.start,
                         "data/Raw/SUKIA_20200204_040000.wav"))
rec21 <- cutWave(rec21, from = 185, to = 186)
writeWave(rec21,
          paste0(
            path.start,
            "data/Processed/templates/rec_20200204_04_05pm.wav"
          ))

# Template 22: SUKIA_20191030_010000.wav (01:05)
rec22 <- readWave(paste0(path.start,
                         "data/Raw/SUKIA_20191030_010000.wav"))
rec22 <- cutWave(rec22, from = 203, to = 204)
writeWave(rec22,
          paste0(
            path.start,
            "data/Processed/templates/rec_20191030_01_05pm.wav"
          ))

# Template 23: SUKIA_20191119_20000.wav
rec23 <- readWave(paste0(path.start,
                         "data/Raw/SUKIA_20191119_200000.wav"))
rec23 <- cutWave(rec23, from = 35, to = 36)
writeWave(rec23,
          paste0(
            path.start,
            "data/Processed/templates/rec_220191119_20_05pm.wav"
          ))

# Template 24: Data/SUKIA_20191020_22000.wav (22:15)
rec24 <- readWave(paste0(path.start,
                         "data/Raw/SUKIA_20191020_220000.wav"))
rec24 <- cutWave(rec24, from = 915, to = 916)
writeWave(rec24,
          paste0(
            path.start,
            "data/Processed/templates/rec_20191020_22_15pm.wav"
          ))

#Template 25: Data/SUKIA_20191113_21000.wav (21:05)
rec25 <- readWave(paste0(path.start,
                         "data/Raw/SUKIA_20191113_210000.wav"))
rec25 <- cutWave(rec25, from = 948, to = 949)
writeWave(rec25,
          paste0(
            path.start,
            "data/Processed/templates/rec_20191113_21_05pm.wav"
          ))

#Template 26: Data/SUKIA_20191220_18000.wav (18:15)
rec26 <- readWave(paste0(path.start,
                         "data/Raw/SUKIA_20191220_180000.wav"))
rec26 <- cutWave(rec26, from = 912, to = 913)
writeWave(rec26,
          paste0(
            path.start,
            "data/Processed/templates/rec_20191220_18_15pm.wav"
          ))
```

Create templates from clipped recordings
```{r}


temp.3 <-
  makeCorTemplate(
    clip = paste0(
      path.start,
      "data/Processed/templates/rec_20190906_21_05pm.wav"
    ),
    name = "T3",
    t.lim = c(0, 0.35),
    frq.lim = c(2.2, 3.1))

temp.5 <-
  makeCorTemplate(
    clip = paste0(
      path.start,
      "data/Processed/templates/rec_20191011_21_05pm.wav"
    ),
    name = "T5",
    t.lim = c(0 , 0.3),
    frq.lim = c(2, 3)) 


temp.8 <-
  makeCorTemplate(
    clip = paste0(
      path.start,
      "data/Processed/templates/rec_20191102_22_05pm.wav"
    ),
    name = "T8",
    t.lim = c(0.4 , 0.7),
    frq.lim = c(2.1, 3)
  )

temp.9 <-
  makeCorTemplate(
    clip = paste0(
      path.start,
      "data/Processed/templates/rec_20191102_22_05pm.wav"
    ),
    name = "T9",
    t.lim = c(0.32 , 0.7),
    frq.lim = c(2.3, 2.7)
  )

temp.10 <-
  makeCorTemplate(
    clip = paste0(
      path.start,
      "data/Processed/templates/rec_20191204_01_05pm.wav"
    ),
    name = "T10",
    t.lim = c(0.1 , 0.4),
    frq.lim = c(2.1, 2.8)
  )

temp.11 <-
  makeCorTemplate(
    clip = paste0(
      path.start,
      "data/Processed/templates/rec_20191219_22_05pm.wav"
    ),
    name = "T11",
    t.lim = c(0.5 , 0.9),
    frq.lim = c(1.8, 3)
  )

temp.12 <-
  makeCorTemplate(
    clip = paste0(
      path.start,
      "data/Processed/templates/rec_20191215_04_15pm.wav"
    ),
    name = "T12",
    t.lim = c(0.7 , 1),
    frq.lim = c(2.1, 2.7)
  )





# temp.15 <-
#   makeCorTemplate(clip = paste0(             path.start,             "data/Processed/templates/rec_20200114_03_15pm.wav"),
#                   name = "T15",
#                   select = "rect")

temp.16 <-
  makeCorTemplate(
    clip = paste0(
      path.start,
      "data/Processed/templates/rec_20191202_20_00pm.wav"
    ),
    name = "T16",
    t.lim = c(0.17 , 0.6),
    frq.lim = c(2.2, 2.8)
  )

temp.17 <-
  makeCorTemplate(
    clip = paste0(
      path.start,
      "data/Processed/templates/rec_20200102_23_15pm.wav"
    ),
    name = "T17",
    t.lim = c(0.54 , 0.83),
    frq.lim = c(2, 3)
  )


temp.19 <-
  makeCorTemplate(
    clip = paste0(
      path.start,
      "data/Processed/templates/rec_20200131_03_15pm.wav"
    ),
    name = "T19",
    t.lim = c(0.32, 0.8),
    frq.lim = c(2, 3)
  )


temp.21 <-
  makeCorTemplate(
    clip = paste0(
      path.start,
      "data/Processed/templates/rec_20200204_04_05pm.wav"
    ),
    name = "T21",
    t.lim = c(0.38 , 0.65),
    frq.lim = c(1.8, 3)
  )

temp.22 <-
  makeCorTemplate(
    clip = paste0(
      path.start,
      "data/Processed/templates/rec_20191030_01_05pm.wav"
    ),
    name = "T22",
    t.lim = c(0.48 , 0.8),
    frq.lim = c(2, 3)
  )

temp.23 <-
  makeCorTemplate(
    clip = paste0(
      path.start,
      "data/Processed/templates/rec_220191119_20_05pm.wav"
    ),
    name = "T23",
    t.lim = c(0.66 , 1),
    frq.lim = c(2.2, 3)
  )

temp.24 <-
  makeCorTemplate(
    clip = paste0(
      path.start,
      "data/Processed/templates/rec_20191020_22_15pm.wav"
    ),
    name = "T24",
    t.lim = c(0.1, 0.5),
    frq.lim = c(2, 3)
  )

temp.25 <-
  makeCorTemplate(
    clip = paste0(
      path.start,
      "data/Processed/templates/rec_20191113_21_05pm.wav"
    ),
    name = "T25",
    t.lim = c(0, 0.5),
    frq.lim = c(2.2, 2.8)
  )

temp.26 <-
  makeCorTemplate(
    clip = paste0(
      path.start,
      "data/Processed/templates/rec_20191220_18_15pm.wav"
    ),
    name = "T26",
    t.lim = c(0.4, 0.8),
    frq.lim = c(2.1, 2.8)
  )

# Crear una lista con todos los templates 

lemur.temp <-
  combineCorTemplates(
   
    temp.3,
    temp.5,
    temp.8
    # temp.9,
    # temp.10,
    # temp.11,
    # temp.12,
    # temp.16,
    # temp.17,
    # temp.19,
    # temp.21,
    # temp.22,
    # temp.23,
    # temp.24,
    # temp.25,
    # temp.26
  )

#plot(lemur.temp)
```

## Survey

Set the cutoff for the survey
```{r}
cutoff <- 0.55
templateCutoff(lemur.temp)[1:length(lemur.temp)]<-rep(cutoff,length(lemur.temp))
```

Quantify A. lemur calls
```{r, message=FALSE, results='hide'}
# Create a vector with filenames and a table to register the number of calls per recording
fnames5 <- dir(paste0(path.start, "data/Processed/Five_Min_Recs/"),
               pattern = "SUKIA")
detect.counts <- data.frame(new.name = fnames5)

for (i in 1:length(fnames5)) {
  # Read in survey recording
  rec.fp <-
    file.path(paste0(path.start, "data/Processed/Five_Min_Recs/"),
              paste0(fnames5[i]))
  
  # Match to template
  cscores <- corMatch(rec.fp, lemur.temp)
  
  # Detect and quantify calls
  cdetects <- findPeaks(cscores)
  detecciones <- getDetections(cdetects)
  if (nrow(detecciones) > 0) {
    cantos <- timeAlign(detecciones)
  } else{
    cantos <- 0
  }
  
  # Add count to table
  if (nrow(detecciones) > 0) {
    detect.counts$count[i] <-
      nrow(cantos)
  } else{
    detect.counts$count[i] <- cantos
  }
} 
i=67
``` 


Read in manually annotated data
```{r}
man.dat <-
  read.csv(
    file = paste0(path.start, "data/Processed/TrainingDat.csv"),
    header = T,
    sep = ","
  )

# add column with new name
# dash after year
man.dat$new.name <- gsub(pattern = "(A_[0-9]{4})" ,
       replacement = "\\1-",
       x = man.dat$file.name)

# dash after month
man.dat$new.name <- gsub(pattern = "(A_[0-9]{4}-[0-9]{2})" ,
                         replacement = "\\1-",
                         x = man.dat$new.name)

# Remove previous directory name
man.dat$new.name <- gsub(pattern = "Data/" ,
                         replacement = "\\1",
                         x = man.dat$new.name)

# Name recordings with end time
for (i in 1:length(man.dat$periodo)) {
  if ("05" %in% unlist(strsplit(x = as.character(man.dat$periodo[i]), split = ":"))[2]) {
    man.dat$new.name[i] <- gsub(pattern = "0000.wav" ,
                                replacement = "0500.wav",
                                x = man.dat$new.name[i])
    
  }
}

for (i in 1:length(man.dat$periodo)) {
  if ("15" %in% unlist(strsplit(x = as.character(man.dat$periodo[i]), split = ":"))[2]) {
    man.dat$new.name[i] <- gsub(pattern = "0000.wav" ,
                                replacement = "2000.wav",
                                x = man.dat$new.name[i])
    
  }
}

# Add time zone
man.dat$new.name <- gsub(pattern = ".wav" ,
                         replacement = "_UTC-6.wav",
                         x = man.dat$new.name)

man.dat <- man.dat[, c(1,10,2:9)]
```

Merge metadata with detection counts
```{r}
surv.dat <- merge(x = man.dat, y = detect.counts, by = "new.name")
```

Plot counts by abundance index
```{r}
ggplot(data = surv.dat, aes(
  x = indice.abun,
  y = count,
  colour = as.factor(indice.abun)
)) +
  geom_jitter (width = 0.2,
               size = 3,
               alpha = 0.7) +
  labs (x = "Indice abundancia", y = "Numero cantos") +
  scale_color_manual(values = wes_palette("Darjeeling1")[c(2, 3, 4, 1)], guide = F) +
  theme_light(base_size = 16)
```